---
title: "Calculate total areas for figures"
author: "Johannes Piipponen"
date: "26 01 2022"
output: html_document
---

# Introduction 

In this document we calculate figure classes. For example, Fig 2 "% of total area" calculated here.
Moreover, Table S1 calculated at the end of this document

# Calculate surface areas for different figure classes
figures.rmd must be read in

```{r}
# function that help to calculate areas for figure classes
f_area <- function(lower_bound, upper_bound) {
  var_low_up <- var_to_examine 
  var_low_up[var_low_up < lower_bound] <- NA
  var_low_up[var_low_up >=  upper_bound] <- NA
  
  # raster of actual grassland area of the class (eg class agb < 25)
  r_of_areas <- cellSize(var_low_up, unit = "km") * mcd_01_5arcmin 
  
  # sum of all cells
  return(sum(values(r_of_areas), na.rm = TRUE))
}

```



# Figure 2a areas (median ab and cc) 

```{r}
var_to_examine <- ab_5y_med/1e3

# Calculate total grassland area
# First, calculate cell sizes, then multiply by the fraction of grassland area that actually exist in each cell
grassland_area_raster <- 
  cellSize(var_to_examine, unit = "km") * mcd_01_5arcmin 
# sum these values
(total_area <- sum(values(grassland_area_raster), na.rm = T)) # 54148698
##(total_area_wrong <- expanse(var_to_examine, unit = "km")) #  98184600 # this computes sum of all cells that has some grassland

summary(ab_5y_med/1e3)
100*f_area(-Inf, 25)/total_area
100*f_area(25, 50)/total_area
100*f_area(50, 100)/total_area
100*f_area(100, 200)/total_area
100*f_area(200, 400)/total_area
100*f_area(400, Inf)/total_area
22 + 12 +20 +20 +20+ 6



# cc quick check --> sligthly different values
var_to_examine <- cc_5y_med
100*f_area(-Inf, 5)/total_area # total area is the same as earlier
100*f_area(5, 10)/total_area
100*f_area(10, 20)/total_area
100*f_area(20, 41)/total_area
100*f_area(41, 82)/total_area
100*f_area(82, Inf)/total_area
22 + 12 + 20 + 20 + 20 + 6


```

# Figure 2b cc trend

```{r}
var_to_examine <-  cc_trend$estimate*15
# total grassland area defined earlier


# modify function f_area a bit (just some cropping as mcd_01_5arcmin as the extent differ due to some na rows we had to remove from lm_rast.tif before regression
f_area_lm <- function(lower_bound, upper_bound) {
  var_low_up <- var_to_examine 
  var_low_up[var_low_up < lower_bound] <- NA
  var_low_up[var_low_up >=  upper_bound] <- NA
  
  # raster of actual grassland area of the class (eg class agb < 25)
  r_of_areas <- cellSize(var_low_up, unit = "km") * 
    crop(mcd_01_5arcmin, var_to_examine)
  
  # sum of all cells
  return(sum(values(r_of_areas), na.rm = TRUE))
}



## all trend classes
100*f_area_lm(-Inf, -20) / total_area
100*f_area_lm(-20, -5) / total_area
100*f_area_lm(-5, 0) / total_area
100*f_area_lm(0, 5) / total_area
100*f_area_lm(5, 20) / total_area
100*f_area_lm(20, Inf) / total_area
8 + 13 + 6 + 4 + 6 + 5 # = 42, so non-significant trend covers the rest, 58%

```


# Figure 2c IV

```{r}
var_to_examine <- ab_iv 


summary(ab_iv)
100*f_area(-Inf, 5)/total_area # total area the same, 54148698 km2
100*f_area(5, 10)/total_area
100*f_area(10, 20)/total_area
100*f_area(20, 30)/total_area
100*f_area(30, 40)/total_area
100*f_area(40, Inf)/total_area
1 + 14 + 44 + 24 + 8 + 9


```



# Figure 2d min to med ratio

```{r}
var_to_examine <- 100 *ab_min_to_med  

summary(100 *ab_min_to_med)
100*f_area(-Inf, 10)/total_area
100*f_area(10, 20)/total_area
100*f_area(20, 40)/total_area
100*f_area(40, 60)/total_area
100*f_area(60, 80)/total_area
100*f_area(80, Inf)/total_area

2 + 3 + 5 + 14 + 47 + 29

```



# Figure 4 RSD maps 

```{r}
# rsd_5y_med
var_to_examine <-  rsd_5y_med
summary(rsd_5y_med)
100*f_area(-Inf, 0.20)/total_area
100*f_area(0.20, 0.65)/total_area
100*f_area(0.65, Inf)/total_area
42 + 28 + 30



# rsd_max
var_to_examine <- rsd_max


# total area sligtly different due to rowmax function
grassland_area_raster_rsd_max <- 
  cellSize(rsd_max, unit = "km") * mcd_01_5arcmin 
# sum these values
(total_area_rsd_max <- sum(values(grassland_area_raster_rsd_max), na.rm = T)) # 53543173
summary(rsd_max)
100*f_area(-Inf, 0.20)/total_area_rsd_max
100*f_area(0.20, 0.65)/total_area_rsd_max
100*f_area(0.65, Inf)/total_area_rsd_max
35 + 26 + 39




# rsd_years_overstocked
var_to_examine <- rsd_years_overstocked
summary(rsd_years_overstocked)

# calculate total area
grassland_area_raster_rsd_overstocked <- 
  cellSize(rsd_years_overstocked, unit = "km") * mcd_01_5arcmin 
## sum these values
(total_area_rsd_overstocked <-
    sum(values(grassland_area_raster_rsd_overstocked), na.rm = T)) # 55429466


100*f_area(-Inf, 0.00001)/total_area_rsd_overstocked #
100*f_area(0.00001, 20)/total_area_rsd_overstocked 
100*f_area(20, 40)/total_area_rsd_overstocked 
100*f_area(40, 60)/total_area_rsd_overstocked
100*f_area(60, 80)/total_area_rsd_overstocked
100*f_area(80, Inf)/total_area_rsd_overstocked
62 + 5 + 3 + 3 + 3+ 24


```


# Figure 5 livestock-only systems

CC in livestock-only grasslands
RSD in livestock-only grasslands

```{r}
# cc_5y_glp --> cc masked to glp areas
var_to_examine <-  cc_5y_glp

# calculate total area
grassland_area_raster_glp <- 
  cellSize(cc_5y_glp, unit = "km") * mcd_01_5arcmin 
## sum these values
(total_area_glp <-
    sum(values(grassland_area_raster_glp), na.rm = T)) # 39205571



hist(cc_5y_glp, xlim= c(0, 100), breaks = 100)
100*f_area(-Inf, 5)/total_area_glp
100*f_area(5, 10)/total_area_glp
100*f_area(10, 21)/total_area_glp
100*f_area(21, 42)/total_area_glp
100*f_area(42, 83)/total_area_glp
100*f_area(83, Inf)/total_area_glp

24 + 12 + 19 + 18 + 21 + 6




# rsd_5y_glp --> cc masked to glp areas
var_to_examine <-  rsd_5y_glp
hist(rsd_5y_glp, xlim= c(0, 1), breaks= 1000)
100*f_area(-Inf, 0.20)/total_area_glp
100*f_area(0.20, 0.65)/total_area_glp
100*f_area(0.65, Inf)/total_area_glp

47 + 29 + 24
```


# Figure 6 CV


```{r}
# ab_cv
var_to_examine <- 100*ab_cv

# calculate total area
grassland_area_raster_ab_cv <- 
  cellSize(100*ab_cv, unit = "km") * mcd_01_5arcmin 
## sum these values
(total_area_ab_cv <-
    sum(values(grassland_area_raster_ab_cv), na.rm = T)) # 54149067


summary(100*ab_cv)
100*f_area(15, 20)/total_area_ab_cv
100*f_area(20, 25)/total_area_ab_cv
100*f_area(25, 30)/total_area_ab_cv
100*f_area(30, 35)/total_area_ab_cv
100*f_area(35, 40)/total_area_ab_cv
100*f_area(40, Inf)/total_area_ab_cv
66 + 20 + 11 + 3 + 0 + 0





# rsd_cv
var_to_examine <- 100*rsd_cv

# calculate total area
grassland_area_raster_rsd_cv <- 
  cellSize(100*rsd_cv, unit = "km") * mcd_01_5arcmin 
## sum these values
(total_area_rsd_cv <-
    sum(values(grassland_area_raster_rsd_cv), na.rm = T)) # 39205571


summary(100*rsd_cv)
100*f_area(15, 20)/total_area_rsd_cv
100*f_area(20, 25)/total_area_rsd_cv
100*f_area(25, 30)/total_area_rsd_cv
100*f_area(30, 35)/total_area_rsd_cv
100*f_area(35, 40)/total_area_rsd_cv
100*f_area(40, Inf)/total_area_rsd_cv
4 + 19 + 39 + 26 + 9 + 3

```


# Supplementary figures

! need to be modified

```{r}

# ISIMIP AGB
var_to_examine <- agb_med_isimip_2006_2010/1e3 # from figures_supplementary.Rmd


# calculate total grassland area for isimip maps
grassland_area_raster_isimip <- 
  cellSize(agb_med_isimip_2006_2010, unit = "km") * mcd_01_5arcmin 
 ## sum these values
(total_area_isimip <-
    sum(values(grassland_area_raster_isimip), na.rm = T)) # 46019190




summary(agb_med_isimip_2006_2010/1e3)
100*f_area(-Inf, 25)/total_area_isimip
100*f_area(25, 50)/total_area_isimip
100*f_area(50, 100)/total_area_isimip
100*f_area(100, 200)/total_area_isimip
100*f_area(200, 400)/total_area_isimip
100*f_area(400, Inf)/total_area_isimip
21 + 9 + 16 + 22+ 25 + 7



# ISIMIP IV AGB
var_to_examine <- agb_iv_isimip_2001_2010 
summary(agb_iv_isimip_2001_2010)
100*f_area(-Inf, 5)/total_area_isimip
100*f_area(5, 10)/total_area_isimip
100*f_area(10, 20)/total_area_isimip
100*f_area(20, 30)/total_area_isimip
100*f_area(30, 40)/total_area_isimip
100*f_area(40, Inf)/total_area_isimip
5 + 23 +45 + 20 + 4 + 3



# ISIMIP CV AGB
var_to_examine <- 100*agb_cv_isimip_2006_2010
summary(100*agb_cv_isimip_2006_2010)
100*f_area(15, 20)/total_area_isimip
100*f_area(20, 25)/total_area_isimip
100*f_area(25, 30)/total_area_isimip
100*f_area(30, 35)/total_area_isimip
100*f_area(35, 40)/total_area_isimip
100*f_area(40, Inf)/total_area_isimip
1 + 6 + 10 + 13 + 12 + 58



# MODIS_AGB / ISIMIP_AGB
var_to_examine <- 100*agb_div
summary(100*agb_div)
100*f_area(0, 50)/total_area_isimip
100*f_area(50, 75)/total_area_isimip
100*f_area(75, 100)/total_area_isimip
100*f_area(100, 125)/total_area_isimip
100*f_area(125, 150)/total_area_isimip
100*f_area(150, Inf)/total_area_isimip
10 + 17 + 25 + 24 + 14 + 10



# RSD calculated with areal weighted animals
var_to_examine <-  crop(rsd_5y_med_aw, mcd_01_5arcmin) # this rsd derived in figures_supplementary_Rmd
summary(rsd_5y_med_aw)
100*f_area(-Inf, 0.20)/total_area # total area as above, 54148698
100*f_area(0.20, 0.65)/total_area
100*f_area(0.65, Inf)/total_area
40 + 30 + 30




```





# Extra calculations 

-total CC and CC in livestock-grazing areas
-total number of animal units (AU) in our study area and livestock-grazing grasslands


```{r}
# cc per pixel (originally per km2 and therefore we multiply with cellsize)
# we also multiply this by the fraction of real grassland area as CC is expressed as AU/km2 which means that if there is grassland in the cell, the density is xx AU/km2
cc_per_pixel <-  cc_5y_med * cellSize(cc_5y_med, unit = "km") * mcd_01_5arcmin
plot(cc_per_pixel)


# total CC (area = IGBP classes 8-10)
global(cc_per_pixel, fun = "sum", na.rm = TRUE ) / 1e6 # 1537 million AU
#sum(values(cc_per_pixel), na.rm = T) / 1e6




# mask cc to livestock-only grasslands
cc_5y_med_glp <- mask(cc_5y_med, glp_5as)
# cc per pixel in livestock-only systems. Again, multiply with the "real" grassland area
cc_per_pixel_glp <- cc_5y_med_glp * cellSize(cc_5y_med_glp, unit = "km") * mcd_01_5arcmin
plot(cc_per_pixel_glp)
#total cc in livestock-only areas
global(cc_per_pixel_glp, fun = "sum", na.rm = TRUE ) / 1e6 # 1117



# cc in glp areas compared to cc in whole study area
global(cc_per_pixel_glp, fun = "sum", na.rm = TRUE ) / 
  global(cc_per_pixel, fun = "sum", na.rm = TRUE )  # 73%


# GLW total numbers
# first, non simulated animal numbers
data_for_simulations <- 
  here("Data", "Output", "data_for_simulations.tif") %>% rast() 


global(data_for_simulations$cattle,
       fun = sum,  na.rm = TRUE ) / 1e6 # 1335 million cows
global(data_for_simulations$sheep,
       fun = sum,  na.rm = TRUE ) / 1e6 # 968 million sheep 
global(data_for_simulations$buffalo,
       fun = sum,  na.rm = TRUE ) / 1e6 #  143 million buffalo 
global(data_for_simulations$horse,
       fun = sum,  na.rm = TRUE ) / 1e6 #  59 million horse 
global(data_for_simulations$goat,
       fun = sum,  na.rm = TRUE ) / 1e6 #  778 million goat 

# this is alltogether something between
1335*0.5 + 968*0.10 + 143 * 0.6 + 59*0.4 + 778*0.1 # 951 m AU when minimum conversion factors are used, see S2.8
# and
1335*1.25 + 968*0.15 + 143 * 0.7 + 59*1.8 + 778*0.15 # 2137 m AU when max animal unit conversion factors are used, see S2.8



## # GLW modelled (and simulated) animal units
# We want to explore total number of animal units in our study area and livestock-grazing grasslands
au_per_km2 <- sim_df %>% 
  dplyr::select(x,y,au_pkm2_med_2010)  %>%
  rasterFromXYZ(., crs = epsg4326) %>%
  rast()

# Areal weighted (aw) product
# sim_aw_animals <- here("Data","Supplementary", "simulation_results_global_n1000_aw_animals.csv") %>% 
#   fread()
# names(sim_aw_animals)

# au_per_km2 <- sim_aw_animals %>% 
#   dplyr::select(x,y,au_pkm2_med_2010)  %>% 
#   rasterFromXYZ(., crs = epsg4326) %>% 
#   rast()




# as unit is per km2, we have to multiply this with pixelarea to get number of au per pixel
au_per_pixel <-  au_per_km2 * cellSize(au_per_km2, unit = "km")

#total AU in our study area
global(au_per_pixel, fun = "sum", na.rm = TRUE ) / 1e6 # 1492 AU



#AU in livestock-only systems
## au per pixel masked in glp areas
au_per_km2_glp <- crop(au_per_km2, glp_5as) %>% 
  mask(., glp_5as)
au_per_pixel_glp <- au_per_km2_glp * cellSize(au_per_km2_glp, unit = "km")
global(au_per_pixel_glp, fun = "sum", na.rm = TRUE ) / 1e6 # 629 million AU  was 671.1269	million AU


# tibble of abs values
tibble(
  cc_total = global(cc_per_pixel, fun = "sum", na.rm = TRUE ) / 1e6,
  cc_total_livestock_grazing = global(cc_per_pixel_glp, fun = "sum", na.rm = TRUE ) / 1e6,
  au_total = global(au_per_pixel, fun = "sum", na.rm = TRUE ) / 1e6,
  au_total_livestock_grazing = global(au_per_pixel_glp, fun = "sum", na.rm = TRUE ) / 1e6) %>% 
  unlist() 

# tibble of relative shares
tibble(
 cc_glp_of_total_cc = global(cc_per_pixel_glp, fun = "sum", na.rm = TRUE ) /
   global(cc_per_pixel, fun = "sum", na.rm = TRUE ),
 au_glp_of_total_au =  global(au_per_pixel_glp, fun = "sum", na.rm = TRUE )/
    global(au_per_pixel, fun = "sum", na.rm = TRUE ),
 potential_used = global(au_per_pixel, fun = "sum", na.rm = TRUE ) /
   global(cc_per_pixel, fun = "sum", na.rm = TRUE ),
 potential_used_glp = global(au_per_pixel_glp, fun = "sum", na.rm = TRUE ) /
   global(cc_per_pixel_glp, fun = "sum", na.rm = TRUE ) ) %>% 
  unlist() 


```



