---
title: "Countries AGB, CC and RSD"
author: "Johannes Piipponen"
date: "26 01 2022"
output: html_document
---


# Introduction 

In this document we do calculations for individual countries. These will be presented both in the main text but also in the Appendix.

1) total grassland area per country
2) total AGB per country 
3) total CC per country
4) how many % of the grassland area is being overgrazed (per country)
5) how many % the grassland area is from the total area of country
6) median RSD per country
7) changes in grassland extent in selected countries
8) total AGB estimates (restricted, unrestricted, ISIMIP based) for Figure S5



# Get data

Packages, sim_df, cc_med and rsd_5y_med data loaded in the file 
figures.Rmd 
Polygons from:
https://www.naturalearthdata.com/downloads/10m-cultural-vectors/10m-admin-0-countries/


*polygons*
```{r}
adm_10m <- here("Data", "Input", "ne_10m_admin_0_countries.shp") %>% 
 read_sf()

library(rmapshaper); library(broom) # other packages from Figures.Rmd
# select only variables needed 
adm_10m <- adm_10m %>% 
  filter(!ADMIN %in% c("Antarctica" )) %>%  #"Greenland"
  mutate(ID = 1:nrow(.)) %>% 
    dplyr::select(ID, SOVEREIGNT, ADMIN) # do not simplify before extracting


# selected countries (countriesthat  Wolf et al used)
mycountries_list <- c("Argentina", "Brazil", "Canada", "Chile", "China", "India", "Kazakhstan", "Mexico", "Russia", "United States of America")

# polygons for selected countries --> use ADMIN for selected countries
mycountries <- subset(adm_10m, adm_10m$ADMIN %in% mycountries_list)
# all countries
countries_list <- adm_10m$ADMIN

# convert to vect --> needed when extracting values
mycountries_extract_polygons <- mycountries %>% 
  as("Spatial") %>%
  vect()

extract_polygons <- adm_10m %>%
  as("Spatial") %>%
  vect() 
 ## add total area of polygons to adm_10m
adm_10m$country_total_area_sqkm <-  expanse(extract_polygons, unit = "km")

```



# 1-5

Althought our maps (resolution 5 arcmin) cover a large fraction of the globe, the total area is smaller in reality. This results in aggregation. Therefore, we need information about the "real" grassland area each of the cells in our 5armin map contains. In other words, aboveground biomass (AGB) and carrying capacity (CC) are density variables and we must calculate AGB and CC in areas where the grasslands actually exist. This information is derived in mcd_01_5arcmin raster (see Figures.Rmd)

*data for the extract*
```{r}
# real grassland area in world. 
grassland_area_world <- cellSize(mcd_01_5arcmin, unit = "km") * mcd_01_5arcmin
# plot(cellSize(cc_5y_med , unit = "km"), main = "cellsizes") 
# plot(mcd_01_5arcmin, main = "fraction of cell that is actually grassland") 
# plot(grassland_area_world, main = "real grassland area")


# AGB and CC
#ab_5y_med %>% plot() # this is density. Interpretation as follows: "if there is grassland in cell, this shows how much biomass exist on those areas". We want total sums; therefore we use values per pixel
agb_5y_perpixel <- ab_5y_med * cellSize(ab_5y_med, unit = "km") * mcd_01_5arcmin
cc_5y_perpixel <- cc_5y_med * cellSize(cc_5y_med, unit = "km") * mcd_01_5arcmin



# overstocked grasslands only (grassland where RSD > 0.65)
rsd_overstocked_world <- rsd_5y_med
rsd_overstocked_world[rsd_overstocked_world < 0.65] <- NA
# test1 <- rsd_overstocked_world; test1[test1>= 0.65] <- 1; plot(test1) # basic picture


 ## areas of overstocked grasslands
overstocked_area_world <- 
  cellSize(rsd_overstocked_world, unit = "km") * mcd_01_5arcmin
plot(overstocked_area_world)





# create brick
countrydata <- c(grassland_area_world, agb_5y_perpixel,
                 cc_5y_perpixel, overstocked_area_world)
names(countrydata) <- c("grassland_area_world", "agb_5y_perpixel",
                        "cc_5y_perpixel", "overstocked_area_world")

```

## extract for every country

*extract*
```{r}
# extract 
countrydata_extracted <- terra::extract(x = countrydata, 
                                        y = extract_polygons,
                                        fun = sum, na.rm = TRUE)




 ## add extracted data to sf object with geometries and plot
adm_10m <- adm_10m %>% 
  mutate(grassarea_sum = countrydata_extracted$grassland_area_world,
         agb_sum = countrydata_extracted$agb_5y_perpixel, 
         cc_sum = countrydata_extracted$cc_5y_perpixel,
         overstocked_areas_sum = countrydata_extracted$overstocked_area_world,
         total_area_sum = countrydata_extracted$total_land_area_world) %>% 
  
  mutate(fraction_overgrazed = 100*overstocked_areas_sum/ grassarea_sum,
         fraction_grassland = 100*grassarea_sum/country_total_area_sqkm) 



 ## plot with tmap (better plot below)
tm_shape(adm_10m, projection = "+proj=robin") +
  tm_polygons(col = c("grassarea_sum", "agb_sum",
                      "overstocked_areas_sum", "fraction_overgrazed"))

```




 ## 6) add rsd median per country to adm_10m 

```{r}

# RSD median
rsd_med_extracted <- terra::extract(x = rsd_5y_med,  y = extract_polygons,
                fun = median, na.rm = TRUE)




adm_10m <- adm_10m %>% 
  mutate(rsd_median = rsd_med_extracted$rsd_med_median) 



# save country specific data as csv
adm_10m_df <- as.data.frame(adm_10m) %>% 
  rename(., total_grassland_area_km2 = grassarea_sum,
         total_agb_in_kg = agb_sum,
         total_cc_in_au = cc_sum,
         total_overgrazed_grassland_area_km2 = overstocked_areas_sum,
         share_of_countrys_grasslands_overgrazed = fraction_overgrazed,
         share_of_grasslands_of_total_land_area_of_country = fraction_grassland,
         median_rsd_value_of_country = rsd_median) %>% 
  dplyr::select(-geometry)


write_csv2(adm_10m_df,
                file = here("Data", "Supplementary", "Provided_supplementary_material",
                            "table_country_specific_data_based_on_median_cc_of_2008_2012.csv"))



# create a table of AGB and CC by country
adm_10m_df_table <- adm_10m_df %>%
  mutate(across(c(country_total_area_sqkm, median_rsd_value_of_country), round, 2))

# create html table
library(kableExtra)
plt_adm_10m_df_table <-
adm_10m_df_table %>%
  kbl(caption = "Tabultaed country-specific results") %>%
  kable_classic(full_width = F, html_font = "Cambria") #%>%
plt_adm_10m_df_table


```


# Plot
 ## function for plotting polygons
 Adapted from figures.Rmd. 
 
```{r}
create_map_polygons <- function(r_index, r_colnames, index_label,index_main_title,
                             colorpal, breakvals,
                             breaknames = NULL,
                             color_midpoint = NULL, tocrs = NA){

  # create tmap object
  index_map <- tm_shape(r_index, projection = "+proj=robin") + 
    tm_polygons(col = r_colnames,
                palette = colorpal,
                breaks = breakvals,
                labels = breaknames,
              title = index_label,
              midpoint = color_midpoint,
              legend.reverse = TRUE) +
    tm_layout(main.title = index_main_title,
              main.title.position = "left",
              main.title.size = 1,
              legend.bg.color = TRUE,
              legend.outside = TRUE,
              frame = FALSE)+
    tm_borders(col = "grey30", lwd = 0.33) 
  
  return (index_map) # Thanks for the function Vili Virkki
}

```
 
 
 
# Figure 5, Figure S6

We want both absolute values and relative shares
For example, total grassland area per country and share of grassland area of country's total land area


```{r}
pal_cc <- scico(n = 6, palette = "bamako", direction = -1)
pal_npp <- scico(n = 6,begin = 0.4, end = 1, palette = "corkO")
pal_total_grassland <- scico(n = 6, palette = "davos", direction = -1)
pal_share_of_max <-  scico(n = 6, palette = "hawaii", begin = 0, end = 1.0, direction = -1)
pal_min_to_med <-  scico(n = 3, palette = "hawaii", begin = 0.1, end = 0.8)

# simplify --> only for plotting
adm_10m_rob <-  st_transform(adm_10m, crs = "ESRI:54030")
adm_10m_simple <- ms_simplify(adm_10m_rob) # this removes +50 rows from adm_10m


# RSD fractions and sums
 ## median rsd per country
pal_rsd2 <- scico(n = 6, palette = "lajolla")


(plt_median_rsd_country <- 
   create_map_polygons(r_index = adm_10m_simple,
                    r_colnames = "rsd_median" ,
                    colorpal = pal_rsd2,
                    breakvals = c(0, 0.10, 0.20, 0.40, 0.65, 0.80, Inf),
                    index_label = "rsd",
                    index_main_title = "Median rsd by country") )



 ## total sum of overgrazed areas per country
(plt_total_overgrazed_area <- 
   create_map_polygons(r_index = adm_10m_simple,
                    r_colnames = "overstocked_areas_sum" ,
                    colorpal = pal_total_grassland,
                    breakvals = c(0, 10000, 1e5, 2e5,5e5, 3e6, Inf),
                    index_label = "km2",
                    index_main_title = "Total overgrazed grassland area by country") )





 ## overgrazed area of country's total grassland area
pal_rsd_years_overst <- scico(n = 6, palette = "hawaii", begin = 0.2,end = 0.7,
                              direction = -1)

(plt_overgrazed_rsd_country <- 
   create_map_polygons(r_index = adm_10m_simple,
                    r_colnames = "fraction_overgrazed",
                    colorpal = pal_rsd_years_overst,
                    breakvals = c(0, 10, 20, 40, 60, 80, 100),
                    index_label = "[%]",
                    index_main_title = "Share of country's grassland area overgrazed"))



(plt_overgrazed_by_country <- tmap_arrange(
  plt_median_rsd_country,
  plt_total_overgrazed_area,
  plt_overgrazed_rsd_country,
  ncol = 2))



tmap_save(plt_overgrazed_by_country,
          here("Figures", "Figure5_overgrazing_by_country.pdf"),
          width = 180, height = 150, units = "mm")







# total grassland area per country
(plt_total_grassland_area <- 
    create_map_polygons(r_index = adm_10m_simple,
                    r_colnames = "grassarea_sum" ,
                    colorpal = pal_total_grassland,
                    breakvals = c(0, 1e5, 2e5,5e5, 2e6, 5e6, Inf),
                    index_label = "km2",
                    index_main_title = "Total grassland area per country") )




# country's grassland area / country's total land area
(plt_share_of_grasslands_of_total_land_area <- 
    create_map_polygons(r_index = adm_10m_simple,
                    r_colnames = "fraction_grassland",
                    colorpal = pal_min_to_med, # pal_share_of_max
                    breakvals = c(0, 10, 20, 40, 60, 80, 100),
                    index_label = "%",
                    index_main_title = "Share of grassland area of total land area by country") )



# total biomass per country
(plt_total_biomass <- 
   create_map_polygons(r_index = adm_10m_simple,
                    r_colnames = "agb_sum" ,
                    colorpal = pal_npp,
                    breakvals = c(0, 1e9, 1e10, 1e11, 5e11, 1.2e12,  Inf), # c(0, 1e6, 5e7, 1e8, 5e8, 1e9,  Inf)
                    index_label = "kg",
                    index_main_title = "Total aboveground biomass (AGB) sum by country") )



(plts_grassland_areas <- 
  tmap_arrange(
  plt_share_of_grasslands_of_total_land_area,
  plt_total_grassland_area,
  plt_total_biomass,
  ncol = 2))

tmap_save(plts_grassland_areas,
           here("Figures", "Supplementary", "Figure_S6bcd_tot_grass_country_and_fraction_country.pdf"),
          width = 180, height = 150, units = "mm")


```







# Varying grassland extent

We used mode value of land cover class in the main text but here we explored how grassland extent changed over 2001-2015

```{r}
rastlist_mcd_year <- function(year) {
  list.files(path = here("Data", "Supplementary"),
             pattern= paste0("MCD_different_extent_", year), 
             full.names = TRUE) }


rastlist_mcd_year("2001") #check 



# function for getting the rasterlist and creating a mosaic
f_mosaic_year_mcd <- function(year) {
  writeRaster(raster(ext = e, crs = epsg4326),
            file=paste0(here("Data", "Supplementary"),"/mosaic_mcd_varying_",
                        year ,".tif"),
            format="GTiff",
            overwrite=TRUE ) 
  
  mosaic_rasters(gdalfile=rastlist_mcd_year(year),
               dst_dataset=paste0(here("Data", "Supplementary"), "/mosaic_mcd_varying_",year, ".tif"),
               of="GTiff")
} 

f_mosaic_year_mcd("2001") #for testing. merges raster for 2001


# run function for all years
sapply(timestep, f_mosaic_year_mcd) # this takes time 1115



# get the merged tiffs
mymergedfiles_mcd_varying <- 
  list.files(path = here("Data", "Supplementary"),
             pattern="mosaic_mcd_varying_",
             full.names=TRUE)

mymergedfiles_mcd_varying
mcd_varying <- rast(mymergedfiles_mcd_varying) 
mcd_varying 
names(mcd_varying) <- c(paste0("mcd", timestep))
plot(mcd_varying)

# aggregate and resample using MCD with 0101010101010101010101 data.
# Then we can find mean grassland extent (potential)  in 5 arc-min raster
# and define some threshold value for "grassland". E.g. aggregated mean < 5% in 5arcmin cell --> not really grassland
tic()
mcd_varying_0_1_agg <- aggregate(mcd_varying, fact = 18, fun = "mean", na.rm = T)
toc() # 41min

tic()
mcd_varying_0_1_5arcmin <- 
  resample(mcd_varying_0_1_agg, 
           rast(here("Data", "Input", "5_Ct_2010_Da.tif")),
           filename = here("Data", "Supplementary", 
                              "mcd_varying_2001_2015_1_or_0_5arcmin.tif"))
toc() # 1 min

mcd_varying_0_1_5arcmin$mcd2015 %>% plot() # send this data to Mika?


############################################
# get the data
mcd_varying_0_1_5arcmin <- 
  here("Data", "Supplementary", "mcd_varying_2001_2015_1_or_0_5arcmin.tif") %>% 
  rast()

mcd_varying_0_1_5arcmin %>% plot()
############################################









# extract total grassland area for selected countries. 
grassland_area_world_15yrs <- cellSize(mcd_varying_0_1_5arcmin, unit = "km") * mcd_varying_0_1_5arcmin
plot(grassland_area_world_15yrs)


mcd_area_15yrs_mycountries <- terra::extract(x = grassland_area_world_15yrs/1e6,
                                 y = mycountries_extract_polygons, fun = sum, na.rm = T)

# no geometries needed, add just the name of SOVEREIGNT to the df
mcd_area_15yrs_mycountries <- mcd_area_15yrs_mycountries %>% 
  mutate(ADMIN = mycountries$ADMIN) %>% 
  dplyr::select(-ID)
names(mcd_area_15yrs_mycountries) <- c(paste0("mcd", timestep), "ADMIN")



 ## test plot
mcd_area_15yrs_mycountries_long <- mcd_area_15yrs_mycountries %>% 
 #  filter(SOVEREIGNT %in% c("Chile", "Argentina", "India", "Mexico")) %>% 
  pivot_longer(cols = mcd2001:mcd2015,
               names_to = "year",
               names_prefix = "mcd",
               values_to = "million_km2" )



  ### plot plot
(p_mcd_mycountries <- mcd_area_15yrs_mycountries_long %>% 
  ggplot(., aes(x = year,
                y = million_km2,
                colour = ADMIN, group = ADMIN)) +
  geom_line(size = 0.5) +
  ggtitle("Total grassland area in selected countries over 2001-2015")+
    
  # Add labels at the end of the line
  geom_text(data = filter(mcd_area_15yrs_mycountries_long, year == "2015"),
            aes(label = ADMIN),
            hjust = 0, nudge_x = 0.1) +
    
  # Allow labels to bleed past the canvas boundaries
  coord_cartesian(clip = 'off') +
    
  # Remove legend & adjust margins to give more space for labels
  theme(plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
        panel.background = element_rect(fill = "white"),
        axis.text.x = element_text(angle = 40, hjust = 1),
        legend.position = 'none',
        plot.margin = margin(0.1, 5.6, 0.1, 0.1, "cm")) ) +
  
  scale_y_continuous(breaks =c(1,2,3,4,5,6,7,8))


ggsave(filename = here("Figures", "Supplementary",
                       "Figure_S6e_total_grassland_area_in_selected_countries_linegraph.pdf"),
              plot = p_mcd_mycountries,
       width = 180, height = 150, units = "mm")



```


# Total CC over 2001-2015 for different countries (use mode mcd value)  --> Kendall trend

Extract carrying capacity (CC) trend for individual countries and calculate trend significance with Kendall's tau



```{r}
# CC
cc_med_per_pixel <- cc_med * cellSize(cc_med, unit = "km") * mcd_01_5arcmin
  # crop(mcd_varying_0_1_5arcmin, cc_med) 

plot(cc_med_per_pixel$cc_med_2015)  ;#plot(cc_5y_perpixel)



 ## extract CC
cc_over_2001_2015 <- terra::extract(x = cc_med_per_pixel, y = extract_polygons,
                                        fun = sum, na.rm = TRUE) #256 rows



  ## combine with adm
adm_10m <- left_join(adm_10m, cc_over_2001_2015, by = "ID")


# calculate significances of countrywise CC trends based on Kendals tau
## function for cor.test()
f_cortest_countries <- function(admin_name) {
  test_kendall <- cor.test(x = 1:length(timestep),
           y = adm_10m %>%
             as.data.frame() %>% 
             filter(ADMIN == paste0(admin_name)) %>%
             dplyr::select(cc_med_2001:cc_med_2015) %>%
             unlist(),
           method = "kendall")
  return(test_kendall %>%  broom::tidy() %>% dplyr::select(estimate, p.value))
}

 ## test
f_cortest_countries("Finland")
 ##  for all countries
df_cortest <- map_df(countries_list, f_cortest_countries) %>% 
  mutate(ID = 1:nrow(.)) 
 # df_cortest %>% filter(p.value > 0.05) %>% nrow() #82 insignificant


 ## combine with geometries
adm_10m_cortest <- left_join(adm_10m, df_cortest, by = "ID") %>% 
  dplyr::select(ID, SOVEREIGNT, ADMIN, cc_med_2001:cc_med_2015, estimate, p.value)

 ## recode so that insignificants get estimate value of -999
adm_10m_cortest <- adm_10m_cortest %>% 
  mutate(estimate.significant = case_when(
    p.value <= 0.05 ~ estimate,
    p.value > 0.05 ~  -999)) %>%  # maybe NA even better
  mutate(estimate.significant = na_if(estimate.significant, -999))




 ## save
cc_over_2001_2015_df <- as.data.frame(adm_10m_cortest) %>%
  dplyr::select(-geometry)
# 
write_csv2(cc_over_2001_2015_df,
                file = here("Data", "Supplementary", "Provided_supplementary_material",
                            "table_total_cc_in_AUs_2001_2015_countries.csv"))



# plot
pal_trend <- scico(n = 6, palette = "roma")


(plt_cc_kendall_country <-
    create_map_polygons(r_index = ms_simplify(adm_10m_cortest),
                    r_colnames = "estimate.significant" ,
                    colorpal = pal_trend,
                    breakvals = c(-1, -0.5,0, 0.5, 1),
                    breaknames = c("< -0.5", "-0.5, 0", "0, 0.5", "> 0.5"),
                    index_label = "AU",
                    index_main_title = "Significant change in CC over 2001-2015 by country [Kendall's tau]") )



tmap_save(plt_cc_kendall_country, filename = here("Figures", "Figure3b_cor_test_kendall_country.pdf"),
          width = 180, height = 150, units = "mm")
```



# 7) MCD -- total grassland area per country for each year 2001-2015 --> here we use varying mcd

Varying land cover is not addressed in the manuscript (only for selected countries in Figure S6e in the Appendix). However, we provide this as supplementary data.

```{r}
mcd_varying_perpixel <-  cellSize(mcd_varying_0_1_5arcmin, unit = "km") * mcd_varying_0_1_5arcmin 
names(mcd_varying_perpixel) <- c(paste0("grassarea", timestep))
 ## extract
mcd_varying_over_2001_2015 <- terra::extract(x = mcd_varying_perpixel, y = extract_polygons,
                                        fun = sum, na.rm = TRUE)
 ## add admin and sovereignt
mcd_varying_over_2001_2015_df <- as.data.frame(mcd_varying_over_2001_2015) %>% 
  mutate(Sovereignt = adm_10m$SOVEREIGNT,
         Admin = adm_10m$ADMIN) %>% 
  dplyr::select(Sovereignt, Admin, grassarea2001:grassarea2015)


write_csv2(mcd_varying_over_2001_2015_df,
                file = here("Data", "Supplementary", "Provided_supplementary_material",
                            "table_varying_grasslandarea_in_km2_2001_2015_countries.csv"))
```



# 8) Comparison of AGB to other global studies Wolf et al. (2021) and Sun et al. (2021)
See Figure S5 in the Appendix

```{r}
ab_med_2006_2010 <- sim_df %>% 
  dplyr::select(x,y,ab_med_2006, ab_med_2007, ab_med_2008, ab_med_2009, ab_med_2010) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast()


agb_med_2006_2010_perpixel <- ab_med_2006_2010 * cellSize(ab_med_2006_2010, unit = "km") * mcd_01_5arcmin


# extract sum of agb for countries
agb_sum_2006_2010_extracted <- terra::extract(x = agb_med_2006_2010_perpixel, 
                                        y = extract_polygons,
                                        fun = sum, na.rm = TRUE)

 # find median of these sums
agb_sum_2006_2010_extracted <- agb_sum_2006_2010_extracted %>% 
  dplyr::select(-ID) %>% 
    mutate(agb_median_of_2006_2010_sums =
           rowMedians(as.matrix(.),na.rm = TRUE))


 # add this median to adm_10m
adm_10m <- adm_10m %>% 
  mutate(agb_median_2006_2010 = agb_sum_2006_2010_extracted$agb_median_of_2006_2010_sums) 



Figure_S5_MODIS_NPP_based_AGB_with_restrictions <- adm_10m %>% 
  dplyr::select(-country_total_area_sqkm, -geometry) %>% 
  filter(ADMIN %in% mycountries_list)






########################################## same for isimip (data from figures.supplementary.Rmd)
# AGB ISIMIP
ab_med_2006_2010_isimip <- sim_isimip %>% 
  dplyr::select(x,y,agb_med_2006,agb_med_2007,agb_med_2008, agb_med_2009,agb_med_2010) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast()

agb_med_2006_2010_perpixel_isimip <- 
  ab_med_2006_2010_isimip * cellSize(ab_med_2006_2010_isimip, unit = "km") * mcd_01_5arcmin


 ## extract sum of agb for countries ISIMIP
agb_sum_2006_2010_extracted_isimip <- terra::extract(x = agb_med_2006_2010_perpixel_isimip, 
                                        y = extract_polygons,
                                        fun = sum, na.rm = TRUE)



 # find median of these sums
agb_sum_2006_2010_extracted_isimip <- agb_sum_2006_2010_extracted_isimip %>% 
  dplyr::select(-ID) %>% 
    mutate(agb_median_of_2006_2010_sums_isimip =
           rowMedians(as.matrix(.),na.rm = TRUE))

 # add this median to adm_10m
adm_10m <- adm_10m %>% 
  mutate(agb_median_2006_2010_isimip = agb_sum_2006_2010_extracted_isimip$agb_median_of_2006_2010_sums_isimip) 




Figure_S5_MODIS_ISIMIP_based_AGB_with_restrictions  <- adm_10m %>% 
  dplyr::select(-country_total_area_sqkm, -agb_median_2006_2010) %>% 
  filter(ADMIN %in% mycountries_list)

```


# extract unrestricted AGB sums by country for Figure S5 (in the Appendix)

```{r}
agb_5arcmin_unrestricted <- 
  here("Data", "Output", "agb_no_simulation_unrestricted_biomass_kg_km2_5arcmin.tif") %>% 
  rast()
names(agb_5arcmin_unrestricted) <- c(paste0("agb_unrest_", timestep))



# select only variables needed
agb_unrestricted_2006_2010 <-
  agb_5arcmin_unrestricted[[c("agb_unrest_2006", "agb_unrest_2007", "agb_unrest_2008",
                                "agb_unrest_2009", "agb_unrest_2010")]]

# crop to match mcd_01_5arcmin
agb_unrestricted_2006_2010 <- crop(agb_unrestricted_2006_2010, mcd_01_5arcmin)

# calculate AGB unrestricted per pixel

agb_unrestricted_2006_2010_perpixel <- agb_unrestricted_2006_2010 * cellSize(agb_unrestricted_2006_2010, unit = "km") * mcd_01_5arcmin


# extract sum of agb for countries
agb_unrestricted_sum_2006_2010_extracted <- terra::extract(x = agb_unrestricted_2006_2010_perpixel, 
                                        y = extract_polygons,
                                        fun = sum, na.rm = TRUE)



 # find median of these sums
agb_unrestricted_sum_2006_2010_extracted <- agb_unrestricted_sum_2006_2010_extracted %>% 
  dplyr::select(-ID) %>% 
    mutate(agb_unres_median_of_2006_2010_sums =
           rowMedians(as.matrix(.),na.rm = TRUE))


 # add this median to adm_10m
adm_10m <- adm_10m %>% 
  mutate(agb_unres_median_2006_2010 = 
           agb_unrestricted_sum_2006_2010_extracted$agb_unres_median_of_2006_2010_sums) 




Figure_S5_MODIS_NPP_based_AGB_unrestricted  <- adm_10m %>% 
  dplyr::select(-country_total_area_sqkm) %>% 
  filter(ADMIN %in% mycountries_list) 
```

 
 