---
title: "Figures"
author: "Johannes"
date: "Jan 2022"
output: html_document
---


# Introduction

Here we plot figures for the main text. However, country-specific figures created in the file
extract_countries.Rmd 
and areas for different figure classes calculated in the file
areas_for_fig_classes.Rmd

Figures are further modified in adobe illustrator. Note that aboveground biomass abbreviated here as AB instead of AGB. 

The files needed in the analysis should be used in following order:
1) creating_TreeCoverMultiplier.Rmd
2) GEE_codes.txt
3) rast_processing_res000449.Rmd
4) simulation.Rmd
5) cc_trend_with_linear_regression.Rmd
6) figures.Rmd
7) extract_countries.Rmd
8) gather_isimip_data.Rmd
9) simulate_isimip_agb.Rmd
10) gather_glw_aw_for_figure_S4.Rmd
11) areas_for_fig_classes.Rmd



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

```{r include = FALSE}
packages <- c("tidyverse", "raster","gdalUtils", "scico","tmap", "tmaptools",
              "sf", "terra", ,"here", "rmapshaper",
              "data.table", "broom", "Rfast", "matrixStats",
              "easypackages")

not_installed <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(not_installed)){install.packages(not_installed)}
library(easypackages) # loads all the libraries
libraries(packages)

# or
# library(tidyverse); library(raster); library(gdalUtils); library(scico); library(tmap); library(tmaptools); library(sf); library(terra); library(here); library(rmapshaper)
# library(data.table); library(broom); library(Rfast); library(matrixStats)
timestep <- 2001:2015
epsg4326 <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

# options, mainly for terra package
terraOptions(tempdir= here("Temp_R"))
terraOptions()
rasterOptions(tmpdir= here("Temp_R"))
rasterOptions()
```

# Prepare data for the figures
## Data

```{r}
# data from simulation.Rmd (assessed for the uncertanties)
sim_df <- here("Data", "Output", "simulation_results_n1000.csv") %>% # 
  fread()
names(sim_df)





## select data and convert to raster (and change neg to na)
# abovegroun biomass ab
ab_med <- sim_df %>% 
  dplyr::select(x,y,contains("ab_med")) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast() # defined in the simulation that ab => 0.1

# carrying capacity cc 
cc_med <- sim_df %>% 
  dplyr::select(x,y,contains("cc_med")) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast() # ab => 0.1

# relative stocking density rsd
rsd_med <- sim_df %>% 
  dplyr::select(x,y,contains("rsd_med_")) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast() # ab => 0.1






#5y med for AB and CC and RSD
# We calculate this as follows: first take medians for every year 2008,2009,2010,2011 and 2012 then take or median of those. 
ab_med_vars <- sim_df %>%  
  dplyr::select(ab_med_2008,ab_med_2009,ab_med_2010, ab_med_2011,ab_med_2012) 

ab_5y_med <- ab_med_vars %>% 
  mutate(ab_med_median = rowMedians(as.matrix(ab_med_vars),na.rm = TRUE)) %>% 
  bind_cols(sim_df %>% dplyr::select(x,y)) %>%  
  dplyr::select(x,y, ab_med_median) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast() 
#writeRaster(ab_5y_med, filename = here("Data", "Output", "AGB_5y_med_simulated_5arcmin.tif"))


#cc 5y med -->  needed for creating classes for ab cc plot, also later for country-specific results
cc_med_vars <- sim_df %>%  
  dplyr::select(cc_med_2008,cc_med_2009,cc_med_2010, cc_med_2011,cc_med_2012) 

cc_5y_med <- cc_med_vars %>% 
  mutate(cc_med_median = rowMedians(as.matrix(cc_med_vars),na.rm = TRUE)) %>% 
  bind_cols(sim_df %>% dplyr::select(x,y)) %>%  
  dplyr::select(x,y, cc_med_median) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast() 
  ## save
#writeRaster(cc_5y_med, filename = here("Data", "Output", "CC_5y_med_simulated_5arcmin.tif"))


# Calculate this similarly as ab: first medians for every year, then median of those again
rsd_med_vars <- sim_df %>%  
  dplyr::select(rsd_med_2008,rsd_med_2009,rsd_med_2010,rsd_med_2011,rsd_med_2012) 


rsd_5y_med <- rsd_med_vars %>% 
  mutate(rsd_med_median = rowMedians(as.matrix(rsd_med_vars),na.rm = TRUE)) %>% 
  bind_cols(sim_df %>% dplyr::select(x,y)) %>%  
  dplyr::select(x,y, rsd_med_median) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast() 








## CV AB and CC --> take median of CVs calculated every year 2008-2012
#¤ ! CV for CC completely similar
ab_cv_vars <- sim_df %>%  
  dplyr::select(ab_cv_2008,ab_cv_2009,ab_cv_2010, ab_cv_2011,ab_cv_2012) 

ab_cv <- ab_cv_vars %>% 
  mutate(ab_cv_average = rowMedians(as.matrix(ab_cv_vars),na.rm = TRUE)) %>% 
  bind_cols(sim_df %>% dplyr::select(x,y)) %>%  
  dplyr::select(x,y, ab_cv_average) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast() 


## CV RSD 
rsd_cv_vars <- sim_df %>%  
  dplyr::select(rsd_cv_2008,rsd_cv_2009,rsd_cv_2010,rsd_cv_2011,rsd_cv_2012) 

rsd_cv <- rsd_cv_vars %>% 
  mutate(rsd_cv_average = rowMedians(as.matrix(rsd_cv_vars),
                                     na.rm = TRUE)) %>% 
  bind_cols(sim_df %>% dplyr::select(x,y)) %>%  
  dplyr::select(x,y, rsd_cv_average) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast() 








# Interannual variability  --> could be calculated while simulating as well
 ## (using Rfast package)
ab_iv <- sim_df %>%
  dplyr::select(contains("ab_med")) %>%
  mutate(iv_ab = 100* as.matrix(.) %>% rowcvs()) %>%
  bind_cols(sim_df %>% dplyr::select(x,y)) %>%  # get the coordinates
  dplyr::select(x,y, iv_ab) %>%
  rasterFromXYZ(., crs = epsg4326) %>%
  rast()
 ## for CC 
cc_iv <- sim_df %>%
  dplyr::select(contains("cc_med")) %>%
  mutate(iv_cc = 100* as.matrix(.) %>% rowcvs()) %>%
  bind_cols(sim_df %>% dplyr::select(x,y)) %>%  # get the coordinates
  dplyr::select(x,y, iv_cc) %>%
  rasterFromXYZ(., crs = epsg4326) %>%
  rast()





## min to median ratio 
### min_to_med ab
ab_vars <- sim_df %>% dplyr::select(contains("ab_med"))

ab_min_to_med <- ab_vars %>% 
  mutate(ab_min_to_med =
           rowMins(as.matrix(ab_vars), value = TRUE) /
           rowMedians(as.matrix(ab_vars),na.rm = TRUE) ) %>% 
   bind_cols(sim_df %>% dplyr::select(x,y)) %>%  
   dplyr::select(x,y, ab_min_to_med) %>% 
   rasterFromXYZ(., crs = epsg4326) %>% 
   rast()

ab_min_to_med %>% plot()


### min_to_med cc 
cc_vars <- sim_df %>% dplyr::select(contains("cc_med"))

cc_min_to_med <- cc_vars %>% 
  mutate(cc_min = rowMins(as.matrix(cc_vars), value = TRUE),
         cc_med = rowMedians(as.matrix(cc_vars),na.rm = TRUE),
         cc_min_to_med = cc_min / cc_med) %>% 
  bind_cols(sim_df %>% dplyr::select(x,y)) %>%  # get the coordinates
  dplyr::select(x,y,cc_min_to_med) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast()

cc_min_to_med %>% plot


### max rsd pixels (same as RSD for min cc pixels)
rsd_vars <- sim_df %>% dplyr::select(contains("rsd_med"))

rsd_max <-rsd_vars %>% 
  mutate(rsd_max = rowMaxs(as.matrix(rsd_vars), value = TRUE)) %>% # max rsd over 2001-2015
  bind_cols(sim_df %>% dplyr::select(x,y)) %>%  # get the coordinates
  dplyr::select(x,y,rsd_max) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast()





# number of years RSD is overstocked
rsd_med_recoded <- sim_df %>% 
  dplyr::select(contains("rsd_med_")) %>% 
  mutate(across(.cols = everything(),
                ~ if_else(.x < 0.65,  true = 0, false = 1)))#change rows that area below 0.65 to zeros

rsd_years_overstocked <- rsd_med_recoded %>% 
  mutate(yrs_overstocked = rowSums2(as.matrix(rsd_med_recoded),
                                    na.rm = TRUE) ) %>% # rowsums give number of years rsd is 1 1 1 (meaning overgrazed) 
  mutate(yrs_overstocked_percent = 100 *yrs_overstocked/length(timestep) ) %>% 
  bind_cols(sim_df %>% dplyr::select(x,y)) %>% 
  dplyr::select(x,y, yrs_overstocked_percent) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast()

qtm(rsd_years_overstocked)
```

# Grassland area in 5arcmin cells

Aggregate and resample mcd 01 data into 5arcmin. This express a fraction of grassland that exist in each 5arcmin cell

```{r}

mcd <- here("Data", "Output", "mcd_mode_0_or_1_res000449.tif") %>%
  rast()

mcd_01_5arcmin <- crop(mcd, cc_med) %>%
  aggregate(., fact = 18, na.rm = T) %>%
  resample(., cc_med,
           filename = here("Data", "Output", "mcd_fraction_of_grassland_in_cell_5arcmin.tif"),
           overwrite = T)


# get the data
mcd_01_5arcmin <- 
  here("Data", "Output", "mcd_fraction_of_grassland_in_cell_5arcmin.tif") %>% 
  rast()
plot(mcd_01_5arcmin) # 0 to 100% of grassland

```





# Get carrying capacity (CC) trend

Trend calculated with linear regression in the file cc_trend_with_linear_regression.Rmd

```{r include = FALSE}

cc_trend <- here("Data", "Output", "lm_rast_5arcmin.tif") %>% rast()

# remove insignificant values
cc_trend[cc_trend$p.value > 0.05] <- NA

cc_trend %>% plot()
qtm(cc_trend$estimate)

```



## Regions and extract
For macroregion data, see: 

Kummu, M., Ward, P. J., Moel, H. de, & Varis, O. (2010). Is physical water scarcity a new phenomenon? Global assessment of water shortage over the last two millennia. Environmental Research Letters, 5(3), 034006. https://doi.org/10.1088/1748-9326/5/3/034006


```{r}
# region data
reg <- here("Data", "Input", "reg_mollw.gpkg") %>%  st_read()
reg_wgs84 <- st_transform(reg,4326) 

reg_wgs84 <- reg_wgs84 %>% 
   mutate(subregion = c("Australia and Oceania", "Central America",
                          "East Asia", "Eastern Europe and Central Asia",
                          "Ice", "South Asia", "South America", "Middle East",
                          "Sub-Saharan Africa", "North Africa", "North America",
                          "Southeast Asia", "Western Europe")) %>% 
  filter(subregion != "Ice")  # remove Ice

# simplify reg_wgs84 for plotting
reg_wgs84_simple <- ms_simplify(reg_wgs84)

# change to terra based vect
reg_wgs84_vect <- vect(as(reg_wgs84, "Spatial")) # 


# extract CC to subregions - weighted average
# we want to give weight to the cells depending on how much grassland area they contain. 
#Original data just 0 = no grassland, 1 = grassland. Thus, aggregated 5arcmin mcd data tells what is the fraction of grassland in each cell.
#weighted average
# first numerator = sum of CC * weights (fraction of grassland area in range [0,1])
cc_regional_numerator <- 
    extract(x = (cc_med * mcd_01_5arcmin),
          y = reg_wgs84_vect,
          fun= sum, na.rm = TRUE, df = TRUE) 

# then denominator = weights summed over subregion
cc_regional_denominator <-
  extract(x = mcd_01_5arcmin,
          y = reg_wgs84_vect,
          fun= sum, na.rm = TRUE, df = TRUE)   

# join data frames
cc_regional <-
  left_join(cc_regional_numerator, cc_regional_denominator, by = "ID")

# divide cc cols by sum_of_weights
cc_regional <- cc_regional %>% 
  mutate(across(everything()), . / mcd) %>% # name mcd here comes when extracting the denominator
  mutate(subregion = reg_wgs84$subregion) %>% 
  dplyr::select(-ID,-mcd)

```




# Change to the Robinson projection  for figures 

Possible to modify so that projection done only when plotting, but then plotting takes a lot of time.
On the other hand, this does take a lot of RAM.

```{r}
# change to the Robinson projection
# ab_rob <- project(ab_med, "+proj=robin", mask = TRUE) #2min for all layers
# cc_rob <- project(cc_med, "+proj=robin", mask = TRUE)
# rsd_rob <-project(rsd_med, "+proj=robin", mask = TRUE)

# ab and cc and rsd 5y medians
ab_5y_rob <- project(ab_5y_med, "+proj=robin", mask = TRUE) # mask not necessarily needed
cc_5y_rob <- project(cc_5y_med, "+proj=robin", mask = TRUE)
rsd_5y_rob <- project(rsd_5y_med, "+proj=robin", mask = TRUE)


# CVs in % --> separate CVs calculated for 2008, 2009.. 2012 and then we took the median
ab_cv_rob <- 100*project(ab_cv, "+proj=robin", mask = TRUE)
#cc_cv_rob <- 100*project(cc_cv, "+proj=robin", mask = TRUE)
rsd_cv_rob <-100*project(rsd_cv, "+proj=robin", mask = TRUE)

# interannual variability in %
ab_iv_rob <- project(ab_iv, "+proj=robin", mask = TRUE)
cc_iv_rob <- project(cc_iv, "+proj=robin", mask = TRUE)

# min_to_med ratio and RSD max
ab_min_to_med_rob <- 100* project(ab_min_to_med, "+proj=robin", mask = TRUE)
cc_min_to_med_rob <- 100* project(cc_min_to_med, "+proj=robin", mask = TRUE)
rsd_max_rob <- project(rsd_max, "+proj=robin", mask = TRUE)

# years overstocked
rsd_years_overstocked_rob <- project(rsd_years_overstocked, "+proj=robin", mask = TRUE)

# lm based trend for whole study period. CC inc/dec this many AU during 2001-2015
cc_trend_rob <- project(cc_trend$estimate*15 , "+proj=robin", mask = TRUE) 
```


# Function for tmap plotting

```{r}
# Draw a map index 
# @param r_index: a raster to be plotted
# @param index_label: label that will be placed as the legend title
# @param colorpal: color palette to be used, ordered from low to high value
# @param breakvals: break values to be drawn in legend
# @param color_midpoint: TRUE if the color scale has a midpoint, NULL by default (no midpoint in color scale)
# @param tocrs: proj4 string of CRS to which the raster is projected if given (by default, no projection is done)
# @return tmap object of the index given
create_index_map <- function(r_index, index_label,index_main_title,
                             colorpal, breakvals,
                             breaknames = NULL,
                             color_midpoint = NULL, tocrs = NA){

  # project to another crs
  if (!is.na(tocrs)){
    r_index <- project(r_index, tocrs, mask = TRUE)
  }

  # create tmap object
  index_map <- tm_shape(r_index) + # possible to add projection = "+proj=robin" (but slow)
    tm_raster(#style = "cont", # draw gradient instead of classified
              palette = colorpal,
              breaks = breakvals,
              labels = breaknames,
              title = index_label,
              midpoint = color_midpoint,
              legend.reverse = TRUE) +
    tm_layout(main.title = index_main_title,
              main.title.position = "left",
              main.title.size = 1,
              legend.bg.color = TRUE,
              legend.outside = TRUE,
              frame = FALSE)+
    tm_shape(reg_wgs84_simple) +
    tm_borders(col = "grey30", lwd = 0.33) 
  
  return (index_map)
  
} #tmaptools::palette_explorer()
```




# Plot Figure 2a-2d (AB and CC in same figure)

CC = AB / intake_forage. Thus intake_forage = AB/CC and we find median of simulated estimates of animal forage consumption. This can be used to place AB and CC values into the same figure as AB = CC * constant (constant is the intake_forage)

```{r}
global((ab_5y_med) / cc_5y_med, fun = "mean", na.rm = TRUE) # intake 4870kg, depends on the sim
# hist(ab_5y_med/1e3);hist(cc_5y_med) # ;hist(cc_5y_med*4.817)
#animals eat 4817kg per year. Thus CC values that represent AB values are # c(0-5, 5-10, 10-21, 21-42, 42-83)
# check CC histogram and compare to those values


# color palettes
pal_cc <- scico(n = 6, palette = "bamako", direction = -1)
pal_cv <-  scico(n = 6, palette = "nuuk", direction = -1)
pal_min_to_med <-  scico(n = 3, palette = "hawaii", begin = 0.1, end = 0.8)
pal_trend <- scico(n = 6, palette = "roma")

# Aboveground biomass (AB) and carrying capacity (CC) between 2008 and 2012
(plt_abcc_5y <- 
  create_index_map(r_index = ab_5y_rob/1e3, # divide to get correct unit, was kg biomass per km2 and we convert to gram per m2
                   index_main_title = "Median aboveground biomass (AGB) and\n carrying capacity (CC) over 2008-2012",
                   index_label = "AB g/m2/yr",
                   colorpal = pal_cc, #pal_ab,
                   breakvals =  c(-Inf,25,50,100,200,400, Inf),
                   breaknames = c(
                     "< 25 (22% of total area)",
                     "25-50 (12% of total area)",
                     "50-100 (20% of total area)",
                     "100-200 (20% of total area)",
                     "200-400 (20% of total area)",
                     "> 400 (6% of total area)")) )
#22 + 12 +20 +20 +20+ 6




# Trend. Multiplied by number of years to get values in meaningful unit: during 15y period, CC increases/decreases xx animal unit
(plt_abcc_trend <-
  create_index_map(r_index = cc_trend_rob, 
                   index_main_title = "Statistically significant change in CC over 2001-2015 [Animal units (AU)/km2/15yrs]",
                  index_label = "AU",
                   colorpal = pal_trend,
                   breakvals = c(-Inf,-20,-5,0,5,20,Inf),
                   breaknames = c("< -20 (8% of total area)",
                                  "-20, -5 (13% of total area)",
                                  "-5, 0 (6% of total area)",
                                  "0, 5 (4% of total area)",
                                  "5, 20 (6% of total area)",
                                   "> 20 (5% of total area)"),
                   color_midpoint = 0))
#8 + 13 + 6 + 4 + 6 + 5  # and the no significant trend covers the rest, 58%




# Interannual variability in AB and CC
(plt_abcc_iv <- 
  create_index_map(r_index = ab_iv_rob ,
                   index_main_title = "Interannual variability in AGB and CC over 2001-215 [%]",
                   index_label = "IV in %",
                   colorpal = pal_cv,
                   breakvals = c(0,5,10,20,30,40,Inf),
                   breaknames = c("< 5 (1% of total area)",
                                  "5-10 (14% of total area)",
                                  "10-20 (44% of total area)",
                                  "20-30 (24% of total area)",
                                  "30-40 (8% of total area)",
                                  "> 40 (9% of total area)"))) 
# 1 + 14 + 44 + 24 + 8 + 9



# Minimum to median ratio for AB and CC
(plt_abcc_min_to_med <- 
  create_index_map(r_index = ab_min_to_med_rob ,
                   index_main_title = "Minimum to median ratio for AB and CC",
                   index_label = "Ratio in %",
                   colorpal = pal_min_to_med,
                   breakvals = c(0, 10, 20, 40, 60, 80, 100),
                   breaknames = c("0-10 (2% of total area)",
                                  "10-20 (3% of total area)",
                                  "20-40 (5% of total area)",
                                  "40-60 (14% of total area)",
                                  "60-80 (47% of total area)",
                                  "80-100 (29% of total area)"))) 

#2 + 3 + 5 + 14 + 47 + 29





## arrange plots
plt_fig2 <- tmap_arrange(
  plt_abcc_5y,
  plt_abcc_trend,
  plt_abcc_iv,
  plt_abcc_min_to_med,
  ncol = 2)

plt_fig2

tmap_save(plt_fig2, here("Figures", "Figure2.pdf"))
```




# Plot Figure 3a (Regional CC line graph)

Trend significances added (calculated later in this srcipt) in adobe illustrator
Tile 3b calculated in the file
extract_figures.Rmd


```{r}
# CC long
cc_reg_long <- cc_regional %>% 
  pivot_longer(cols = cc_med_2001:cc_med_2015,
               names_to = "year",
               names_prefix = "cc_med_",
               values_to = "AU_km2" )



# plot CC
(p_cc_reg <- cc_reg_long %>% 
  ggplot(., aes(x = year,
                y = AU_km2,
                colour = subregion, group = subregion)) +
  geom_line(size = 0.5) +
 # geom_point(aes(color = subregion))+
    
  ggtitle("Regional Carrying Capacity (CC)")+
    
  # Add labels at the end of the line
  geom_text(data = filter(cc_reg_long, year == "2015"),
            aes(label = subregion),
            hjust = 0, nudge_x = 0.1) +
    
  # Allow labels to bleed past the canvas boundaries
  coord_cartesian(clip = 'off') +
    
  # Remove legend & adjust margins to give more space for labels
  theme(plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
        panel.background = element_rect(fill = "white"),
        axis.text.x = element_text(angle = 40, hjust = 1),
        legend.position = 'none',
        plot.margin = margin(0.1, 5.6, 0.1, 0.1, "cm")) )
   


p_cc_reg

ggsave(filename = here("Figures", "Figure3a.pdf"),
              plot = p_cc_reg)

```





# Plot Figure 4: RSD


```{r}
pal_rsd <- scico(n = 3, palette = "lajolla")

# rsd calculated based on simualated median of cc2008-cc2012
(plt_rsd_med5 <- 
  create_index_map(r_index = rsd_5y_rob,
                   index_main_title = "Relative stocking density (RSD) based on median carrying capacity (CC) over 2008-2012 [%]",
                   index_label = "RSD in %",
                   colorpal = pal_rsd,
                   breakvals = c(-Inf, 0.2, 0.65, Inf),
                   breaknames = c("<20%, low pressure (42% of total area)",
                                  "20-65%, medium pressure (28% of total area)",
                                  ">65%, overstocked (30% of total area)")))
# 42 + 28 + 30



# RSD calculated with min CC values (for every pixel, we calculated min CC)
# this is the same as max simulated RSD values 
(plt_rsd_max <- 
  create_index_map(r_index = rsd_max_rob,
                   index_main_title = "RSD based on minimum CC values over 2001-2015 [%]",
                   index_label = "RSD in %",
                   colorpal = pal_rsd,
                   breakvals = c(-Inf, 0.2, 0.65, Inf),
                   breaknames = c("<20%, low pressure (35% of total area)",
                                  "20-65%, medium pressure (26% of total area)",
                                  ">65%, overstocked (39% of total area)")))

# 35 + 26 + 39



# years when RSD in the class overstocked  #palette_explorer()  alternatives
pal_rsd_years_overst <- scico(n = 6, palette = "hawaii", begin = 0.2,end = 0.7,
                              direction = -1)


(plt_rsd_years_overstocked <- 
  create_index_map(r_index = rsd_years_overstocked_rob,
                   index_main_title = "Share of years 2001-2015 when RSD is in the class overstocked [%]",
                   index_label = "%",
                   colorpal = pal_rsd_years_overst,
                   breakvals =  c(-Inf,0.0001,20, 40, 60, 80, 100),
                   breaknames = c("0 (62% of total area)",
                                  "0-20 (5% of total area)",
                                  "20-40 (3% of total area)",
                                  "40-60 (3% of total area)",
                                  "60-80 (3% of total area)",
                                  "80-100 (24% of total area)") ))

#62 + 5 + 3 + 3 + 3+ 24


## arrange plots
plt_rsd <- tmap_arrange(
  plt_rsd_med5,
  plt_rsd_max,
  plt_rsd_years_overstocked,
  ncol = 2)

plt_rsd


tmap_save(plt_rsd, here("Figures", "Figure4.pdf"))

```


# Figure 5 plotted in the file extract_countries.Rmd
# Livestock-only grasslands, Figure 6


Unfortunately, we cannot share data that divides grasslands to livestock-grazing grasslands. However, the documentation can be found from Robinson et al (2011) and data asked from authors. Here we use classes
LGY (Livestock only, hyper-arid),
LGA (Livestock only, arid and semi-arid),
LGH (Livestock only, humid and sub-humid) and 
LGT (Livestock only, temperate and tropical highland) 
to separate livestock-only systems from mixed (or industrial) production systems. Same as classes 1-4

The glp version used is Ruminant_prod_systemsv5

```{r}
mypath <- "" # change
glp <- rast(paste0(mypath, "GLP/GLP/Ruminat_prod_systemsv5/ruminant_production_systems.tif"))

plot(glp)
# we need only classes 1-4
glp[glp>4] <- NA ; plot(glp)
# change all values to 1
glp[glp>1] <- 1; plot(glp)


# aggregare to 5arcsec, crop and mask to our study area
glp_5as <- aggregate(glp, fact =  10, na.rm = TRUE) %>% 
  crop(., ab_med$ab_med_2010) %>% 
  mask(., ab_med$ab_med_2010)

plot(glp);plot(glp_5as);plot(ab_med$ab_med_2010)

# mask CC and RSD to livestock-only grasslands and plot
# glp denotes global livestock production systems
cc_5y_glp <- cc_5y_med %>% 
  mask(., glp_5as)

rsd_5y_glp<- rsd_5y_med %>% 
  mask(., glp_5as)


# change to rob proj 
cc_5y_rob_glp <- project(cc_5y_glp, "+proj=robin", mask = TRUE)
rsd_5y_rob_glp <- project(rsd_5y_glp, "+proj=robin", mask = TRUE)


# cc_livestock_only
plot(cc_5y_rob_glp)
(plt_cc_5y_glp <- 
  create_index_map(r_index = cc_5y_rob_glp,
                   index_main_title = "Median carrying capacity (CC) over 2008-2012 on the livestock-grazing grasslands",
                   index_label = "CC AU/km2/yr",
                   colorpal = pal_cc,
                   breakvals =  c(0,5,10,21,42,83,Inf),
                   breaknames = c("< 5 (24% of total area)",
                                  "5-10 (12% of total area)",
                                  "10-21 (19% of total area)",
                                  "21-42 (18% of total area)",
                                  "42-83 (21% of total area)",
                                  "> 83 (6% of total area)"))) 
# 24 + 12 + 19 + 18 + 21 + 6


# rsd_livestock_only
(plt_rsd_med5_glp <- 
  create_index_map(r_index = rsd_5y_rob_glp,
                   index_main_title = "Relative stocking density (RSD) based on median carrying capacity (CC) over 2008-2012 on the livestock-grazing grasslands",
                   index_label = "RSD in %",
                   colorpal = pal_rsd,
                   breakvals = c(-Inf, 0.2, 0.65, Inf),
                   breaknames = 
                     c("<20%, low pressure (47% of total area)",
                       "20-65%, medium pressure (29% of total area)",
                       ">65%, overstocked (24% of total area)")))
# 47 + 29 + 24

## arrange plots
(plt_livestock_only <- tmap_arrange(
  plt_cc_5y_glp,
  plt_rsd_med5_glp,
  ncol = 1))

tmap_save(plt_livestock_only, here("Figures", "Figure6.pdf"))
```


# Plot Figure 7: uncertainty measured with coefficient of variation (CV) 

```{r}
pal_cv <-  scico(n = 6, palette = "nuuk", direction = -1)

# ab_cv_rob %>% hist(breaks = 40, xlim = c(15,40), main = "agb_cc_cv")
# rsd_cv_rob%>% hist(breaks = 500, xlim = c(10,70), main = "rsd_cv")


# AGB and CC CV
(plt_ab_cv<-
  create_index_map(r_index = ab_cv_rob ,
                   index_main_title = "Uncertainty related to aboveground biomass (AGB) and carrying capacity (CC)\n measured with coefficient of variation (CV)",
                   index_label = "CV in %",
                   colorpal = pal_cv,
                   breakvals = c(15,20,25,30,35,40,Inf),
                   breaknames = c("15-20 (60% of total area)",
                                  "20-25 (20% of total area)",
                                  "25-30 (13% of total area)",
                                  "30-35 (5% of total area)",
                                  "35-40 (1% of total area)",
                                  "> 40 (1% of total area)") ))
# 60 + 20+ 13 + 5+ 1 + 1


# RSD CV
(plt_rsd_cv <-
  create_index_map(r_index = rsd_cv_rob ,
                   index_main_title = "Uncertainty related to relative stocking density (RSD)\n measured with coefficient of variation (CV)",
                   index_label = "CV in %",
                   colorpal = pal_cv,
                   breakvals = c(15,20,25,30,35,40,Inf),
                   breaknames = c("15-20 (3% of total area)",
                                  "20-25 (21% of total area)",
                                  "25-30 (38% of total area)",
                                  "30-35 (23% of total area)",
                                  "35-40 (10% of total area)",
                                  "> 40 (5% of total area)") ))

# 3 +21 + 38 + 23+ 10 + 5


## arrange plots
plt_cv <- tmap_arrange(
  plt_ab_cv,
  plt_rsd_cv,
  ncol = 1)

plt_cv


# save
tmap_save(plt_cv, here("Figures", "Figure7.pdf"))
```








# Significances of regional CC trends (Figure 3a) based on Kendalls tau

These could be calculated more elegantly as done for individual countries in the file
extract_countries.Rmd
---> significances added to line graph in adobe illustrator

```{r}
f_cortest_region <- function(name_subregion) {
  cor.test(1:length(timestep), # x value for the test
           cc_regional %>%
             filter(subregion == paste0(name_subregion)) %>% 
             dplyr::select( -subregion) %>% 
             unlist(), # y value
           method = "kendall")
}

f_cortest_region("Australia and Oceania") # -0.18, pval 0.38
f_cortest_region("South America") # -0.10, pval 0.63
f_cortest_region("Middle East") # -0.22 pval 0.28
f_cortest_region("Sub-Saharan Africa") #0.01 pval 1
f_cortest_region("North America") #-0.30 pval 0.14
f_cortest_region("North Africa") # 0.52 pval 0.006 ***
f_cortest_region("South Asia") # -0.64 pval 0.0005 ***
f_cortest_region("Central America") # -0.60 pval 0.001 ***
f_cortest_region("East Asia") # -0.31 pval 0.11
f_cortest_region("Eastern Europe and Central Asia") # -0.49 pval 0.011 **
f_cortest_region("Southeast Asia") # -0.62 pval 0.001 ***
f_cortest_region("Western Europe") # -0.77 pval 0.00001 ***

```





