---
title: "Supplementary figures"
author: "Johannes"
date: "01-01-2022"
output: html_document
---

# Introduction
Here we create figures for the Appendix.
However, country specific maps produced in the file
extract_countries.Rmd


```{r include = FALSE}
library(tidyverse); library(raster); library(gdalUtils); library(scico); library(tmap); library(tmaptools); library(sf); library(terra); library(here)
library(data.table); library(broom); library(Rfast); library(matrixStats)
timestep <- 2001:2015
epsg4326 <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
e <- extent(-180,180,-90,90)
```


# Treecover and Net Primary Productivity (NPP) everywhere

Gether files downloaded from Google Earth Engine and combine


```{r}
# treecover. Get files downloaded from Google Earth Engine (GEE)
supplement_files_tc <- list.files(path = here("Data", "Supplementary"),
                           pattern="2000_2015_mean_tree_canopy_cover", full.names=TRUE) 
# NPP. Also from GEE
supplement_files_npp <- list.files(path = here("Data", "Supplementary"),
                           pattern="2000_2015_mean_npp", full.names=TRUE) 

# empty template raster we will fill soon --> note that we use raster package here
writeRaster(raster(ext = e, crs = epsg4326),
            filename = here("Data", "Supplementary",
                            "mosaic_2000_2015_mean_tree_canopy_cover.tif"), ,
            format="GTiff", overwrite=TRUE ) 

# this fills the file created above
mosaic_rasters(gdalfile = supplement_files_tc,
                 dst_dataset = here("Data", "Supplementary", "mosaic_2000_2015_mean_tree_canopy_cover.tif"),
                 of = "GTiff")

# same for the NPP
# empty template raster we will fill soon --> note that we use raster package here
writeRaster(raster(ext = e, crs = epsg4326),
            filename = here("Data", "Supplementary",
                            "mosaic_2000_2015_mean_npp.tif"), ,
            format="GTiff", overwrite=TRUE ) 

# this fills the file created above
mosaic_rasters(gdalfile = supplement_files_npp,
                 dst_dataset = here("Data", "Supplementary", "mosaic_2000_2015_mean_npp.tif"),
                 of = "GTiff")



#### get the merged files 
supplement_tc <- here("Data", "Supplementary", "mosaic_2000_2015_mean_tree_canopy_cover.tif") %>% 
  rast()

supplement_npp <- here("Data", "Supplementary",
                       "mosaic_2000_2015_mean_npp.tif") %>% 
  rast()



## aggregate and resample to 5as resolution. Use cattle GLW raster as template
template_cattle <- here("Data", "Input", "5_Ct_2010_Da.tif") %>% 
  rast()

supplement_tc_agg <- supplement_tc %>% 
  aggregate(., fact = 18, na.rm = TRUE) 


supplement_tc_5arcmin <- supplement_tc_agg %>% 
  resample(., template_cattle,
           filename = here("Data", "Supplementary", "supplementary_treecover_5arcmin.tif"))

supplement_tc_5arcmin %>% plot()
  
# same for npp
supplement_npp_agg <- supplement_npp %>% 
  aggregate(., fact = 18, na.rm = TRUE) 


supplement_npp_5arcmin <- supplement_npp_agg %>% 
  resample(., template_cattle,
           filename = here("Data", "Supplementary",
                           "supplementary_npp_5arcmin.tif"))

supplement_npp_5arcmin %>% plot()

#### get the files 
supplement_npp_5arcmin <- here("Data", "Supplementary",
                           "supplementary_npp_5arcmin.tif") %>% 
  rast()

supplement_tc_5arcmin <-here("Data", "Supplementary", "supplementary_treecover_5arcmin.tif") %>% 
  rast()

```


# Fraction of grassland in larger (5 arc-minutes) cells

```{r}
# get the data created in the file Figures.Rmd 
mcd_01_5arcmin <- 
  here("Data", "Output", "mcd_fraction_of_grassland_in_cell_5arcmin.tif") %>% 
  rast()
plot(mcd_01_5arcmin)

```



# A function for plotting and regions 
Similar function than in the file
Figures.Rmd

```{r}
# Draw a map index 
# @param r_index: a raster to be plotted
# @param index_label: label that will be placed as the legend title
# @param colorpal: color palette to be used, ordered from low to high value
# @param breakvals: break values to be drawn in legend
# @param color_midpoint: TRUE if the color scale has a midpoint, NULL by default (no midpoint in color scale)
# @param tocrs: proj4 string of CRS to which the raster is projected if given (by default, no projection is done)
# @return tmap object of the index given
create_index_map <- function(r_index, index_label,index_main_title,
                             colorpal, breakvals,
                             breaknames = NULL,
                             color_midpoint = NULL, tocrs = NA){

  # project to another crs
  if (!is.na(tocrs)){
    r_index <- project(r_index, tocrs, mask = TRUE)
  }

  # create tmap object
  index_map <- tm_shape(r_index,
                        raster.downsample = TRUE) + 
    tm_raster(#style = "cont", # draw gradient instead of classified
              palette = colorpal,
              breaks = breakvals,
              labels = breaknames,
              title = index_label,
              midpoint = color_midpoint,
              legend.reverse = TRUE) +
    tm_layout(main.title = index_main_title,
              main.title.position = "left",
              main.title.size = 1,
              legend.bg.color = TRUE,
              legend.outside = TRUE,
              frame = FALSE)+
    tm_shape(reg_wgs84) +
    tm_borders(col = "grey30", lwd = 0.5) 
  
  return (index_map)
  
} #tmaptools::palette_explorer()



reg <- here("Data", "Input", "reg_mollw.gpkg") %>%  st_read()
reg_wgs84 <- st_transform(reg,4326) 

reg_wgs84 <- reg_wgs84 %>% 
   mutate(subregion = c("Australia and Oceania", "Central America",
                          "East Asia", "Eastern Europe and Central Asia",
                          "Ice", "South Asia", "South America", "Middle East",
                          "Sub-Saharan Africa", "North Africa", "North America",
                          "Southeast Asia", "Western Europe")) %>% 
  filter(subregion != "Ice")  # remove Ice

# change to terra based vect
reg_wgs84_vect <- vect(as(reg_wgs84, "Spatial")) # 
```

# Figure S2, S3 and S6a


- NPP, gC/m2/yr
- Treecover, percent
- fraction of grassland in larger 5armin cells (plt_fraction_of_grassland)


Originally NPP is kg*C/m2 and the scaling factor is 0.0001. Therefore we divide it by 10 to it in the correct unit of gC/m2/yr

```{r}
# NPP 5arcmin, TC 5arcmin
#change to rob
supplement_npp_5arcmin_rob <- project(supplement_npp_5arcmin, "+proj=robin",
                                      mask = TRUE)

supplement_tc_5arcmin_rob <- project(supplement_tc_5arcmin, "+proj=robin",
                                      mask = TRUE)

pal_npp <- scico(n = 5,begin = 0.6, end = 1, palette = "corkO")
pal_tc <-  scico(n = 6, palette = "bamako", direction = -1)

(plt_npp_mean <- 
  create_index_map(r_index = supplement_npp_5arcmin_rob/10,
                   index_main_title = "Average NPP during the period 2001-2015",
                   index_label = "NPP, gC/m2/yr",
                   colorpal = pal_npp,
                   breakvals =  c(-Inf,100,200,500,1000, Inf),
                   breaknames = c("< 100", "100-200", "200-500","500-1000","> 1000")) )

(plt_tc_mean <- 
  create_index_map(r_index = supplement_tc_5arcmin_rob,
                   index_main_title = "Average tree canopy cover during the period 2001-2015",
                   index_label = "Tree canopy cover, %",
                   colorpal = pal_tc,
                   breakvals =  c(-Inf,5,10,20,30,60,Inf),
                   breaknames = c("< 5", "5-10","10-20","20-30", "30-60", "> 60")) )


# save
tmap_save(plt_npp_mean, here("Figures", "Supplementary", "Figure_S2_average_npp_2001-2015.pdf"),
          width = 180, height = 150, units = "mm")

tmap_save(plt_tc_mean, here("Figures", "Supplementary",
                            "Figure_S3_average_tc_2001-2015.pdf"),
          width = 180, height = 150, units = "mm")




# fraction of grassland in 5arcmin cells
pal_min_to_med <-  scico(n = 3, palette = "hawaii", begin = 0.1, end = 0.8)

supplement_mcd_01_5arcmin_rob <- project(mcd_01_5arcmin*100, "+proj=robin",
                                      mask = TRUE)
 ## remove NA for plotting
supplement_mcd_01_5arcmin_rob[supplement_mcd_01_5arcmin_rob == 0] <- NA

 ## plot
(plt_fraction_of_grassland <- 
    create_index_map(r_index = supplement_mcd_01_5arcmin_rob,
                   index_main_title = "Share of grassland in aggregated 5 arc-minutes cells",
                   index_label = "[%]",
                   colorpal = pal_min_to_med,
                   breakvals =  c(0,10,20,40,60,80, 100),
                   breaknames = c("0-10", "10-20","20-40","40-60","60-80", "80-100")) )


tmap_save(plt_fraction_of_grassland,
          here("Figures", "Supplementary", "Figure_S6a_share_of_grassland_in_5arcmin_cell.pdf"),
          width = 180, height = 150, units = "mm")
```

# Figure S4. Animal densities and RSD derived using DA and Aw methods of GLW

In the main text, we calculated everything using dasymetric (DA) method of Gridded Livestock of the World (GLW). See the documentation in their website. However, for a comparison, we also created a GLW map and RSD map using areal-weighted (Aw) method of GLW. For these Aw method, we ran the same simulation.Rmd script but used Aw animals as input

Note that Figure S4c is the same as the RSD figure presented in the main text (Figure S4a)

```{r}
pal_glw <- scico(n = 6, palette = "lajolla", direction = 1)
# Dasymmetric method (Da) -- this is used everywhere in the main text
 ## Figure S4c is the same as RSD figure presented in the main text
sim_df <- here("Data", "Output", "simulation_results_n1000.csv") %>% # 
  fread()
names(sim_df)


au_per_km2_da <- sim_df %>% 
  dplyr::select(x,y,au_pkm2_med_2010)  %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast()

hist(au_per_km2_da, xlim = c(0,100), breaks = 100)



# change to rob for plotting
au_per_km2_da_rob <- project(au_per_km2_da, "+proj=robin", mask = TRUE)



(plt_glw_simulated_da <- 
  create_index_map(r_index = au_per_km2_da_rob,
                   index_main_title = "Gridded Livestock of the World (GLW) modelled livestock distribution based on simulated animal aunit (AU) conversion factors [AU/km2]. Animal distribution based on dasymetric (DA) method",
                   index_label = "AU/km2",
                   colorpal = pal_glw,
                   breakvals = c(0,1,10,20,50,100,Inf),
                   breaknames = c("< 1",
                                  "1-10",
                                  "10-20",
                                  "20-50",
                                  "50-100",
                                  "> 100")))


tmap_save(plt_glw_simulated_da, here("Figures", "Supplementary",
                            "Figure_S4a_glw_simulated_au_per_km2.pdf"),
          width = 180, height = 150, units = "mm")















# areal weighted (Aw) animals
sim_aw_animals <- here("Data","Supplementary", "simulation_results_global_n1000_aw_animals.csv") %>% 
  fread()
names(sim_aw_animals)

au_per_km2_aw <- sim_aw_animals %>% 
  dplyr::select(x,y,au_pkm2_med_2010)  %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast()

# change to rob for plotting
au_per_km2_aw_rob <- project(au_per_km2_aw, "+proj=robin", mask = TRUE)


#plot GLW aw

(plt_glw_simulated_aw <- 
  create_index_map(r_index = au_per_km2_aw_rob,
                   index_main_title = "Gridded Livestock of the World (GLW) modelled livestock distribution based on simulated animal aunit (AU) conversion factors [AU/km2]. Animal distribution based on areal_weighted (Aw) method",
                   index_label = "AU/km2",
                   colorpal = pal_glw,
                   breakvals = c(0,1,10,20,50,100,Inf),
                   breaknames = c("< 1",
                                  "1-10",
                                  "10-20",
                                  "20-50",
                                  "50-100",
                                  "> 100")))


tmap_save(plt_glw_simulated_aw, here("Figures", "Supplementary",
                            "Figure_S4b_glw_simulated_au_per_km2_aw.pdf"),
          width = 180, height = 150, units = "mm")





# RSD using aw animals as input
rsd_med_vars_aw <- sim_aw_animals %>%  
  dplyr::select(rsd_med_2008,rsd_med_2009,rsd_med_2010,rsd_med_2011,rsd_med_2012) 


rsd_5y_med_aw <- rsd_med_vars_aw %>% 
  mutate(rsd_med_median_aw = rowMedians(as.matrix(rsd_med_vars_aw),na.rm = TRUE)) %>% 
  bind_cols(sim_aw_animals %>% dplyr::select(x,y)) %>%  
  dplyr::select(x,y, rsd_med_median_aw) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast() 



# proj
rsd_5y_med_aw_rob <- project(rsd_5y_med_aw, "+proj=robin", mask = TRUE)

# plot glw
pal_rsd <- scico(n = 3, palette = "lajolla")

# rsd calculated based on simualated median of cc2008-cc2012
(plt_rsd_med5_aw <- 
  create_index_map(r_index = rsd_5y_med_aw_rob,
                   index_main_title = "Relative stocking density (RSD) based on median carrying capacity (CC) over 2008-2012 [%]. Animal distribution based on areal_weighted (Aw) method",
                   index_label = "RSD in %",
                   colorpal = pal_rsd,
                   breakvals = c(-Inf, 0.2, 0.65, Inf),
                   breaknames = c("<20%, low pressure (% of total area)",
                                  "20-65%, medium pressure (% of total area)",
                                  ">65%, overstocked (% of total area)")))



tmap_save(plt_rsd_med5_aw, here("Figures", "Supplementary",
                            "Figure_S4d_rsd_5y_med_simulated_aw.pdf"),
          width = 180, height = 150, units = "mm")



# calculate total animal numbers in our study area (Aw animals)
au_per_pixel_aw <- 
  au_per_km2_aw * cellSize(au_per_km2_aw, unit = "km")

#global(au_per_pixel_aw, fun = "sum", na.rm = TRUE ) / 1e6 # 1478	 million AU (Aw method) 
```



# Figure S7

Compare CC_trend to CC_med
--> how many % of potential animal units (AU) has been lost during 2001-2015?
--> must be cropped first as NA rows were dropped for cc_trend calculations



```{r}
cc_trend # from figures.Rmd
cc_5y_med # from figures.Rmd
ext(cc_5y_med); ext(cc_trend)
cc_5y_med_trendcrop <- crop(cc_5y_med, cc_trend)
comparison_percent <- 100 * (cc_trend$estimate*15) / cc_5y_med_trendcrop
# reproject
comparison_percent_rob <- project(comparison_percent, 
                                  "+proj=robin", mask = TRUE)
         

pal_trend2 <- scico(n = 7, palette = "roma")

(plt_comp_cctrend_cc_med <-
  create_index_map(r_index = comparison_percent_rob,
                   index_main_title = "Ratio of CC trend estimate (multiplied by the number of study years, 15) compared to median CC over 2008-2012 [%]",
                  index_label = "%",
                  colorpal = pal_trend2,
                  breakvals = c(-Inf, -50, -25,-10,10,25,50, Inf),
                   breaknames = c("< -50",
                                  "-50, -25",
                                  "-25, -10",
                                  "-10, 10",
                                  "10, 25",
                                  "25,50",
                                   "> 50"),
                   color_midpoint = 0))


tmap_save(plt_comp_cctrend_cc_med, here("Figures", "Supplementary",
                                        "Figure_S7_ratio_of_cc_trend_to_ccmed.pdf"),
          width = 180, height = 150, units = "mm")
```

# sensitivity plots

```{r}
sense <- here("Data", "Output", "sensitivity_results_2010_n1000_260122.csv") %>% 
  fread()

sense_rob <- rasterFromXYZ(sense, crs = epsg4326) %>% 
  rast() %>% 
  project(., "+proj=robin", mask = TRUE)

names(sense_rob) <- c("NPP", "NPP_temperature",
                      "NPP_temperature_treecover", "NPP_temperature_treecover_animal_units")


pal_cv <-  scico(n = 6, palette = "nuuk", direction = -1)

# CV sensitivity
(plt_sensitivity2010<-
  create_index_map(r_index = 100*sense_rob ,
                   index_main_title = "Model sensitivity to different input parameters in 2010 measured with coefficient of variation (CV)",
                   index_label = "CV in %",
                   colorpal = pal_cv,
                   breakvals = c(-Inf,10,15,20,30,40,Inf) ))


tmap_save(plt_sensitivity2010, here("Figures", "Supplementary", "Figure_S8_sensitivity.pdf"),
          width = 180, height = 150, units = "mm")
```






# Figure S9 - ISIMIP derived aboveground biomass (AGB)

ISIMIP data gathered in the file
gather_isimip_data.Rmd
and simulated in the file
simulate_isimip_agb.Rmd

We'll calculate AGB for 5 year median (years 2006-2010) to make comparison with MODIS NPP derived AGB easier
(those are the compatible years)

```{r}
sim_isimip <- here("Data", "Supplementary", "isimip_agb_1000sim.csv") %>% 
  fread()
names(sim_isimip) # med and cv for every year 2000-2010


# AGB ISIMIP
ab_med_vars_isimip_2006_2010 <- sim_isimip %>%  
  dplyr::select(agb_med_2006,agb_med_2007,agb_med_2008, agb_med_2009,agb_med_2010) 

agb_med_isimip_2006_2010 <- ab_med_vars_isimip_2006_2010 %>% 
  mutate(agb_med_median_isimip = 
           rowMedians(as.matrix(ab_med_vars_isimip_2006_2010),na.rm = TRUE)) %>% 
  bind_cols(sim_isimip %>% dplyr::select(x,y)) %>%  
  dplyr::select(x,y, agb_med_median_isimip) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast() 


# Interannual variability for AGB. Compatible years 2001-2010
agb_iv_isimip_2001_2010 <- sim_isimip %>%
  dplyr::select(contains("agb_med")) %>%
  mutate(iv_agb_isimip = 100* as.matrix(.) %>% rowcvs()) %>%
  bind_cols(sim_isimip %>% dplyr::select(x,y)) %>%  # get the coordinates
  dplyr::select(x,y, iv_agb_isimip) %>%
  rasterFromXYZ(., crs = epsg4326) %>%
  rast()


# CV AGB
agb_cv_vars_isimip_2006_2010 <- sim_isimip %>%  
  dplyr::select(agb_cv_2006,agb_cv_2007,agb_cv_2008, agb_cv_2009,agb_cv_2010) 

agb_cv_isimip_2006_2010 <- agb_cv_vars_isimip_2006_2010 %>% 
  mutate(agb_cv_median =
           rowMedians(as.matrix(agb_cv_vars_isimip_2006_2010),na.rm = TRUE)) %>% 
  bind_cols(sim_isimip %>% dplyr::select(x,y)) %>%  
  dplyr::select(x,y, agb_cv_median) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast() 


# project
agb_med_isimip_2006_2010_rob <- 
  project(agb_med_isimip_2006_2010, "+proj=robin", mask = TRUE)

agb_iv_isimip_2001_2010_rob <- 
  project(agb_iv_isimip_2001_2010, "+proj=robin", mask = TRUE)

agb_cv_isimip_2006_2010_rob <- 100*
  project(agb_cv_isimip_2006_2010, "+proj=robin", mask = TRUE)


# plot
pal_cc <- scico(n = 6, palette = "bamako", direction = -1)
pal_cv <-  scico(n = 6, palette = "nuuk", direction = -1)




## AGB ISIMIP
(plt_agb_med_isimip_2006_2010 <- 
  create_index_map(r_index = agb_med_isimip_2006_2010_rob/1e3, # divide to get correct unit, was kg biomass per km2 and we convert to gram per m2
                   index_main_title = "ISIMIP-based median aboveground biomass (AGB) over 2006-2010",
                   index_label = "AGB g/m2/yr",
                   colorpal = pal_cc, #pal_ab,
                   breakvals =  c(-Inf,25,50,100,200,400, Inf),
                   breaknames = c(
                     "< 25 (% of total area)",
                     "25-50 (%)",
                     "50-100 (%)",
                     "100-200 (%)",
                     "200-400 (%)",
                     "> 400 (%)")) ) # areas_for_fig_classes.Rmd
# Total areas
 # ISIMPI AGB > MODIS AGB when AGB over 200 g m2 yr
 # ISIMPI AGB slightly smaller than MODIS AGB when AGB 25 -100 g m2 yr
 # ! modelled NPP data larger/potentially overestimated in areas with highest NPP values as stated by Irisarri



## IV ISIMIP
(plt_agb_iv_isimip <- 
  create_index_map(r_index = agb_iv_isimip_2001_2010_rob ,
                   index_main_title = "Interannual variability in ISIMIP-based AGB over 2001-2010 [%]",
                   index_label = "IV in %",
                   colorpal = pal_cv,
                   breakvals = c(0,5,10,20,30,40,Inf),
                   breaknames = c("< 5 (% of total area)",
                                  "5-10 (%)",
                                  "10-20 (%)",
                                  "20-30 (%)",
                                  "30-40 (%)",
                                  "> 40 (%)"))) #areas_for_fig_classes.Rmd: 
# Total areas --> different time period
 # ISIMIP has generally smaller variabilities. However, time period is much shorter



## CV AGB ISIMIP
(plt_agb_cv_isimip_2006_2010 <-
  create_index_map(r_index = agb_cv_isimip_2006_2010_rob ,
                   index_main_title = "Uncertainty related to ISIMIP-based aboveground biomass (AGB) measured with coefficient of variation (CV)",
                   index_label = "CV in %",
                   colorpal = pal_cv,
                   breakvals = c(15,20,25,30,35,40,Inf),
                   breaknames = c("15-20 (% of total area)",
                                  "20-25 (%)",
                                  "25-30 (%)",
                                  "30-35 (%)",
                                  "35-40 (%)",
                                  "> 40 (%)") )) #areas_for_fig_classes.Rmd
# Total area
 # CV ISIMIP AGB > CV MODIS AGB. Makes sense as we assumed MODIS NPP varies 7%. Moreover, ISIMIP NPP models include climate forcings --> more uncertain?

```


# Compare ISIMIP and MODIS derived AGB maps

```{r}
# MODIS over 2006-2010 (compatible years with ISIMIP)
agb_med_vars_2006_2010 <- sim_df %>%  # this data from figures.Rmd
  dplyr::select(ab_med_2006,ab_med_2007,ab_med_2008, ab_med_2009,ab_med_2010) 

agb_med_MODIS_2006_2010 <- agb_med_vars_2006_2010 %>% 
  mutate(agb_med_median_MODIS =
           rowMedians(as.matrix(agb_med_vars_2006_2010),na.rm = TRUE)) %>% 
  bind_cols(sim_df %>% dplyr::select(x,y)) %>%  
  dplyr::select(x,y, agb_med_median_MODIS) %>% 
  rasterFromXYZ(., crs = epsg4326) %>% 
  rast() 


# Divide MODIS / ISIMIP
agb_div <- agb_med_MODIS_2006_2010 / agb_med_isimip_2006_2010
agb_div_rob <-  100*project(agb_div, "+proj=robin", mask = TRUE)

agb_div %>% hist()
summary(agb_div)


# plot
pal_trend <- scico(n = 6, palette = "roma")

(plt_division_modis_isimip <-
  create_index_map(r_index = agb_div_rob, 
                   index_main_title = "MODIS derived AGB / ISIMIP derived AGB",
                  index_label = "-",
                   colorpal = pal_trend,
                   breakvals = c(0,50,75,100,125,150,Inf),
                   breaknames = c("0-50% (of total area)",
                                  "50-75% (%)",
                                  "75-100% (%)",
                                  "100-125% (%)",
                                  "125-150% (%)",
                                  "> 150% (%)"),
                   color_midpoint = 0)) #areas_for_fig_classes.Rmd


## arrange plots
plt_isimip <- tmap_arrange(
  plt_agb_med_isimip_2006_2010,
  plt_agb_iv_isimip,
  plt_agb_cv_isimip_2006_2010,
  plt_division_modis_isimip,
  ncol = 2)

plt_isimip

tmap_save(plt_isimip, here("Figures", "Supplementary", "Figure_S9_isimip_vs_modis.pdf"),
          width = 180, height = 150, units = "mm")
```

# Some plots for Figure 1 (flowchart)

```{r}
# plot original slope
slope <- here("Data", "Input", "dtm_slope_merit.dem_m_250m_s0..0cm_2018_v1.0.tif") %>%
  rast()

# aggregate and change to percents. Slopes originally degree * 100 but we need slope% instead 
slope_agg_fact40_percent <- aggregate(slope, fact = 40, na.rm = TRUE) %>% 
  app(., function (x) {100 * tan((x/100)*pi/180  )})



slope_rob <- project(slope_agg_fact40_percent , "+proj=robin", mask = TRUE) 
plot(slope_rob)

pal_slope <- scico(n = 4, palette = "bilbao", begin = 0.1, end = 0.8)



(plt_slope <- 
  create_index_map(r_index = slope_rob, 
                   index_main_title = "Terrain slope",
                   index_label = "%",
                   breakvals =  c(0,10, 30,60,Inf),
                   breaknames = c("0-10", "10-30", "30-60", ">60"),
                   colorpal = pal_slope))

tm_shape(slope_rob) +
  tm_raster(palette = pal_slope,
            breaks = c(0,10, 30,60,Inf),
            labels = c("0-10", "10-30", "30-60", ">60"))
  
  
  

# 
tmap_save(plt_slope, filename = here("Figures", "Supplementary", "slopes_percent_agg40.pdf"),
          width = 180, height = 150, units = "mm")
```
