---
title: "ISIMIP2a"
author: "Johannes Piipponen"
date: "9 12 2021"
output: html_document
---

# Introduction
In this file we gather ISIMIP2a modelled NPP data. See Appendix S2.10

```{r include = FALSE}
library(tidyverse); library(raster); library(gdalUtils); library(scico);
library(sf); library(terra); library(data.table); library(here); library(tictoc)
# Create valid PROJ4 string for EPSG 4326
epsg4326 <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" 
# extents for templaterasters
e <- extent(-180,180,-90,90) 
# timestep isimi
timestep_isimip <- 2000:2010
# options, mainly for terra package
terraOptions(tempdir= here("Temp_R"))
terraOptions()
rasterOptions(tmpdir= here("Temp_R"))
rasterOptions()
wopt_options <- list(gdal = c("BIGTIFF=YES",
                "PREDICTOR = 2",
                "NUM_THREADS = ALL_CPUS"))



data_for_simulations <- here("Data", "Output", "data_for_simulations.tif") %>% 
  rast()
data_for_simulations <- data_for_simulations$npp2015

months <- rep(seq(0,10), each = 12)
```




# get and modify data 1 (caraib)
Note that watch_nobc scenario has only years 1971-2001 so we won't use that
princeton_nobc in for 1971-2012 but we use it only until 2010

"Unit originally C m-2 s-1, and it indicates the average npp in kg per m2 per second of that month. So if you want to have annual total npp in g C m-2, you will need to average(12 monthly npp values) * 1000 g per kg * 86400 second per day * 365 days = g C m-2, where “average(12 monthly npp values) “ is the annual mean npp in kg C m-2 s-1. "


! We want the unit g C m2 yr !


```{r}
rastlist_isimip_caraib <- 
  list.files(path = here("Data", "Supplementary", "ISIMIP"),
             pattern= paste0("caraib"), 
             full.names = TRUE) 

rastlist_isimip_caraib

#conver to raster 
caraib1 <- rastlist_isimip_caraib[1] %>% rast() # gswp3
caraib2 <- rastlist_isimip_caraib[2] %>% rast() # princeton
caraib3 <- rastlist_isimip_caraib[3] %>% rast() # watch-wfdei


# select only months that represent years 2000-2010
caraib1 <- subset(caraib1, c(349:480)) # only years 2000-2010
caraib2 <- subset(caraib2, c(349:480))
caraib3 <- subset(caraib3, c(349:480))


# 1
# yearly averages (multiply with seconds and days)
r <- stack(caraib1 * 365 * 86400)  
all_years <- stack()
for (n in 0:10) {
index <- which(months == n)
annual_mean <- mean(r[[index]], na.rm =T)
all_years <- stack(all_years, annual_mean)}

caraib1_yr <- rast(all_years)
names(caraib1_yr) <- paste0("caraib1_", timestep_isimip)



# 2
r <- stack(caraib2 *  365 * 86400) 
all_years <- stack()
for (n in 0:10) {
index <- which(months == n)
annual_mean <- mean(r[[index]], na.rm =T)
all_years <- stack(all_years, annual_mean)}
caraib2_yr <- rast(all_years)
names(caraib2_yr) <- paste0("caraib2_", timestep_isimip)



# 3
r <- stack(caraib3 *  365 * 86400) 
all_years <- stack()
for (n in 0:10) {
index <- which(months == n)
annual_mean <- mean(r[[index]], na.rm =T)
all_years <- stack(all_years, annual_mean)}
caraib3_yr <- rast(all_years)
names(caraib3_yr) <- paste0("caraib3_", timestep_isimip)


# combine rasters
caraib_all <- c(caraib1_yr, caraib2_yr, caraib3_yr)
names(caraib_all)
plot(caraib_all)
# disaggregate and mask when all ISIMIP rasters combined
```



# get and modify data 2 (lpjml)
```{r}
rastlist_isimip_lpjml <- 
  list.files(path = here("Data", "Supplementary", "ISIMIP"),
             pattern= paste0("lpjml"), 
             full.names = TRUE) 

rastlist_isimip_lpjml # can use only 1-3



#conver to raster 
lpjml1 <- rastlist_isimip_lpjml[1] %>% rast() # gswp3
lpjml2 <- rastlist_isimip_lpjml[2] %>% rast()  # princeton
lpjml3 <- rastlist_isimip_lpjml[3] %>% rast()  # watch-wfdei


# select only months that represent years 2000-2010
lpjml1 <- subset(lpjml1, c(349:480)) # only years 2000-2010
lpjml2 <- subset(lpjml2, c(349:480))
lpjml3 <- subset(lpjml3, c(349:480))


# same things than above
# 1
r <- stack(lpjml1 *  365 * 86400) 
all_years <- stack()
for (n in 0:10) {
index <- which(months == n)
annual_mean <- mean(r[[index]], na.rm =T)
all_years <- stack(all_years, annual_mean)}
lpjml1_yr <- rast(all_years)
names(lpjml1_yr) <- paste0("lpjml1_", timestep_isimip)



# 2
r <- stack(lpjml2 *  365 * 86400) 
all_years <- stack()
for (n in 0:10) {
index <- which(months == n)
annual_mean <- mean(r[[index]], na.rm =T)
all_years <- stack(all_years, annual_mean)}
lpjml2_yr <- rast(all_years)
names(lpjml2_yr) <- paste0("lpjml2_", timestep_isimip)



# 3
r <- stack(lpjml3 *  365 * 86400) 
all_years <- stack()
for (n in 0:10) {
index <- which(months == n)
annual_mean <- mean(r[[index]], na.rm =T)
all_years <- stack(all_years, annual_mean)}
lpjml3_yr <- rast(all_years)
names(lpjml3_yr) <- paste0("lpjml3_", timestep_isimip)


# combine rasters
lpjml_all <- c(lpjml1_yr, lpjml2_yr, lpjml3_yr)
names(lpjml_all)
plot(lpjml_all)
```

# get and modify data 3 (dlem)

```{r}
rastlist_isimip_dlem <- 
  list.files(path = here("Data", "Supplementary", "ISIMIP"),
             pattern= paste0("dlem"), 
             full.names = TRUE) 

rastlist_isimip_dlem

dlem1 <- rastlist_isimip_dlem[1] %>% rast()
dlem2 <- rastlist_isimip_dlem[2] %>% rast()
dlem3 <- rastlist_isimip_dlem[3] %>% rast()

# select only months that represent years 2000-2010
dlem1 <- subset(dlem1, c(349:480)) # only years 2000-2010
dlem2 <- subset(dlem2, c(349:480))
dlem3 <- subset(dlem3, c(349:480))


# same things than above
# 1
r <- stack(dlem1 *  365 * 86400) 
all_years <- stack()
for (n in 0:10) {
index <- which(months == n)
annual_mean <- mean(r[[index]], na.rm =T)
all_years <- stack(all_years, annual_mean)}
dlem1_yr <- rast(all_years)
names(dlem1_yr) <- paste0("dlem1_", timestep_isimip)
# dlem1_dis_cropmask <- disaggregate(dlem1_yr, fact = 6) %>% 
#   crop(., ab_5y_med ) %>% 
#   mask(., ab_5y_med )


# 2
r <- stack(dlem2 *  365 * 86400) 
all_years <- stack()
for (n in 0:10) {
index <- which(months == n)
annual_mean <- mean(r[[index]], na.rm =T)
all_years <- stack(all_years, annual_mean)}
dlem2_yr <- rast(all_years)
names(dlem2_yr) <- paste0("dlem2_", timestep_isimip)
# dlem2_dis_cropmask <- disaggregate(dlem2_yr, fact = 6) %>% 
#   crop(., ab_5y_med ) %>% 
#   mask(., ab_5y_med )


# 3
r <- stack(dlem3 *  365 * 86400) 
all_years <- stack()
for (n in 0:10) {
index <- which(months == n)
annual_mean <- mean(r[[index]], na.rm =T)
all_years <- stack(all_years, annual_mean)}
dlem3_yr <- rast(all_years)
names(dlem3_yr) <- paste0("dlem3_", timestep_isimip)
# dlem3_dis_cropmask <- disaggregate(dlem3_yr, fact = 6) %>% 
#   crop(., ab_5y_med ) %>% 
#   mask(., ab_5y_med )



# combine rasters
dlem_all <- c(dlem1_yr, dlem2_yr, dlem3_yr)
names(dlem_all)
plot(dlem_all)
```

# get and modify data 4 (orchidee)

```{r}
rastlist_isimip_orchidee <- 
  list.files(path = here("Data", "Supplementary", "ISIMIP"),
             pattern= paste0("orchidee"), 
             full.names = TRUE) 

rastlist_isimip_orchidee

orchidee1 <- rastlist_isimip_orchidee[1] %>% rast()
orchidee2 <- rastlist_isimip_orchidee[2] %>% rast()
orchidee3 <- rastlist_isimip_orchidee[3] %>% rast()


# select only months that represent years 2000-2010
orchidee1 <- subset(orchidee1, c(349:480)) 
orchidee2 <- subset(orchidee2, c(349:480))
orchidee3 <- subset(orchidee3, c(349:480))



# same things than above
# 1
r <- stack(orchidee1 *  365 * 86400) 
all_years <- stack()
for (n in 0:10) {
index <- which(months == n)
annual_mean <- mean(r[[index]], na.rm =T)
all_years <- stack(all_years, annual_mean)}
orchidee1_yr <- rast(all_years)
names(orchidee1_yr) <- paste0("orchidee1_", timestep_isimip)



# 2
r <- stack(orchidee2 *  365 * 86400) 
all_years <- stack()
for (n in 0:10) {
index <- which(months == n)
annual_mean <- mean(r[[index]], na.rm =T)
all_years <- stack(all_years, annual_mean)}
orchidee2_yr <- rast(all_years)
names(orchidee2_yr) <- paste0("orchidee2_", timestep_isimip)



# 3
r <- stack(orchidee3 *  365 * 86400) 
all_years <- stack()
for (n in 0:10) {
index <- which(months == n)
annual_mean <- mean(r[[index]], na.rm =T)
all_years <- stack(all_years, annual_mean)}
orchidee3_yr <- rast(all_years)
names(orchidee3_yr) <- paste0("orchidee3_", timestep_isimip)




# combine rasters
orchidee_all <- c(orchidee1_yr, orchidee2_yr, orchidee3_yr)
names(orchidee_all)
plot(orchidee_all)
```


# Combine all rasters, disaggregate and mask
132 layers = 11 years * 3 climate forcing * 4 models 

```{r}
isimip_all_kg_C_m2 <- c(caraib_all, lpjml_all, dlem_all, orchidee_all)
names(isimip_all)

# convert to kg C km2 and remove values smaller than 1
isimip_all_kg_C_km2 <- isimip_all_kg_C_m2 * 1e6
isimip_all_kg_C_km2[isimip_all_kg_C_km2 < 1] <- NA
summary(isimip_all_kg_C_km2)

isimip_all_disag_cropmask <- disaggregate(isimip_all_kg_C_km2, fact = 6) %>% 
  crop(., data_for_simulations) %>%
  mask(., data_for_simulations)

# save as csv
isimip_all_disag_cropmask_df <-
  as.data.frame(isimip_all_disag_cropmask, xy = T) %>% 
  as.data.table()

fwrite(isimip_all_disag_cropmask_df,
       here("Data", "Supplementary", "isimip_npp_kgC_km2_yr.csv"))




# select only year 2000
isimip_2000 <- terra::subset(isimip_all_disag_cropmask,
                             grep('_2000', names(isimip_all),
                                              value = T))
isimip_2000 %>% names()
isimip_2000 # here we find extreme range for uncertainty assessment
isimip_2000 %>% plot()


# find extremes for each cell- Probably best to convert to df
isimip_2000_df <- as.data.frame(isimip_2000, xy = TRUE)

isimip_2000_vars <- isimip_2000_df %>%
  dplyr::select(contains("2000"))

isimip_range_2000 <- isimip_2000_vars %>% 
  mutate(min2000 =rowMins(as.matrix(isimip_2000_vars), value = TRUE),
         max2000 = rowMaxs(as.matrix(isimip_2000_vars), value = TRUE))


```

